### 2025-04-08 manuscript/index.qmd
improved introduction

---
title: "Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation "
author:  
  - name: Erdem M. Karaköylü
    orcid: 0000-0002-6156-1720
    corresponding: true
    email: erdemk@protonmail.com
    roles:
      - Investigation
      - Software
      - Visualization
    affiliations:
      - Consultant
  - name: Susanne E. Craig
    orcid: 0000-0002-8963-0951
    corresponding: true
    email: susanne.e.craig@nasa.gov
    roles:
      - Investigation
      - Project administration
    affiliations:
      - NASA
abstract: Placeholder
date: last-modified
bibliography: references.bib
---

### Introduction

#### Historical Context of Chlorophyll Algorithms
Satellite ocean color observations have long been fundamental for monitoring marine ecosystems, as they enable the global estimation of chlorophyll‑a—a key indicator of phytoplankton biomass and ocean productivity. Early empirical algorithms, notably developed by O’Reilly et al. [@oreilly1998; @oreilly2000], established the $OCx$ family (where $x$ denotes the number of bands used) of polynomial regression models. These models relate blue-to-green reflectance ratios (after log‑transformation) to in situ chlorophyll‑a, employing either straight band ratios (BR) or maximum band ratios (MBR)—the latter selecting the highest available blue-to-green ratio for any given observation as input to a high‑order polynomial. These formulations have served as the operational foundation for chlorophyll‑a products across a broad range of satellite ocean color sensors—from the pioneering Coastal Zone Color Scanner (CZCS) through SeaWiFS, MODIS, and MERIS to more recent missions—offering a straightforward and robust approach for Case‑1 waters. However, their performance is more limited in optically complex Case‑2 waters and remains sensitive to atmospheric correction errors.

Subsequent refinements were introduced to address these deficiencies. For example, Hu et al. [@hu2012novel] proposed a Color Index (CI) formulation that employs a band‑difference approach to reduce sensitivity to residual atmospheric errors and instrument noise, with further improvements enhancing inter‑sensor agreement [@hu2019improving]. The increasing availability of calibration data (e.g., [@Valente2015]) and ongoing algorithmic improvements have led to the development of additional variants of the $OCx$ algorithms—specifically, the OC5 and OC6 formulations. O’Reilly and Werdell [@oreilly2019] maintain that OC5 extends the spectral basis by incorporating the 412 nm band, thereby exploiting its strong signal in clear, oligotrophic waters, while OC6 replaces the traditional denominator with the mean of the 555 and 670 nm reflectances, improving the dynamic range at low chlorophyll concentrations. In total, [@oreilly2019] propose 65 versions of BR/MBR $OCx$ type algorithms for 25 sensors—on average, two or more variants per sensor. With this arsenal, it is hoped, researchers are better equipped to address the wide array of bio‑optical environments encountered in global ocean color applications.

#### Limitations of Existing Approaches
Unfortunately, the development process for traditional ocean color algorithms is founded on a statistical fallacy. Frequentism—the classical approach—assumes that observations are random samples drawn from the studied phenomena. These samples' probabilities are  considered the observed frequencies that capture the the true - albeit noised up and potentially biased - behavior of the system. Under this framework, if a phenomenon is hypothesized to be represented by a model $M$ and data $$D$ is collected, then the model’s representativeness is evaluated solely through the likelihood $p(D∣M)$, which measures the probability of observing the data given that $$M$ is true. The fallacy arises from the mistaken belief that, with sufficiently large amounts of data, one can simply "flip" the likelihood to obtain $p(M∣D)$, thereby yielding a probabilistic statement about $M$. Compounding this error is the imposition of numerous unverified assumptions—disguised as objectivity—that lead frequentists to set the parameters of $M$ by maximizing $p(D∣M)$, under the false premise that this provides the best possible representation of the model. In essence, this fallacy leads to a confusion between the sampling probability $p(D∣M)$ and the inferential probability $p(M∣D)$; the latter  when correctly obtained  quantifies the plausibility of the hypothesis represented by $M$ - precisely what we seek. The historical dimensions of this frequentist illusion are fascinating in their own right, and we invite interested readers to consult works such as McGrayne [@mcgrayne] and Clayton [@clayton].

Bayesianism, Frequentism's archrival, provides the framoework to establish the plausibility of $M$ using observations through the reinterpretation of how probability should be construed. As elaborated extensively by Jaynes [@jaynes2003probability], and more recently by practitioners like Gelman [@gelman2013bayesian], and McElreath [@mcelreath2015statistical], probability is best understood as a statement of our knowledge rather than as an inherent property of the data. the Bayesian perspective treats the data, once observed, as fixed and the model parameters as latent random variables the uncertainty of which  can—and should—be explicitly modeled via prior distributions. Prior distributions are probabilistic constructs that help workers encode prior knowledge. Priors can represents knowledge elicited from subject matter experts, assumptions held by the modeler or a complete absence thereof.

This conceptual framework dissolves the false dichotomy between empirical and semi‑analytical approaches. Both are, in essence, different hypotheses or approximations about how ocean color data are generated; they merely provide alternative expressions of the likelihood 
$p(Chl | Rrs)$ to be updated in light of new data. By adopting a Bayesian paradigm, our approach integrates prior knowledge with observed data to yield a full posterior predictive distribution—offering not only improved predictive performance but also principled uncertainty quantification. In doing so, it circumvents the need for the proliferation of narrowly tuned, sensor-specific variants (65 versions for 25 sensors, as documented by O’Reilly and Werdell [@OREILLY2019]) that are symptomatic of ad hoc calibration processes. Detailed methodological comparisons and additional model refinements are provided in the Supplementary Material.

#### Motivation for Bayesian Tree-Based Models
To address these limitations, we explore Bayesian tree-based modeling – specifically, Bayesian Additive Regression Trees (BART) – as a novel approach for ocean color chlorophyll retrieval. BART offers several advantages for this application:

Direct Bayesian inference: As a Bayesian ensemble method, BART directly targets the posterior distribution $p(Chl | Rrs)$ rather than a purely empirical fit. The model inherently provides a probabilistic mapping from reflectance to chlorophyll. Rather than producing a single best-fit value, BART returns a posterior mean and credible interval for each prediction, naturally quantifying predictive uncertainty.

Uncertainty quantification: The BART framework includes uncertainty in its predictions by design. Every estimated chlorophyll value comes with a confidence range (e.g., 95% credible interval) reflecting model uncertainty and observational noise. This is a key improvement over deterministic algorithms, enabling more reliable use of the data for scientific and operational purposes where understanding uncertainty is crucial e.g. for forecasting trends.

Interpretability: Being a sum-of-trees model, BART can be interrogated to understand the influence of each input variable. Tools such as partial dependence plots allow us to visualize the relationship between each spectral band and the predicted chlorophyll, marginalizing over other bands. In contrast to high-order polynomial coefficients, these tree-based partial dependencies are intuitively interpretable. For instance, one can observe how changes in a particular waveband’s reflectance (holding others average) drive the chlorophyll estimate, revealing any thresholds or non-linear responses.

BART’s modeling approach is non-parametric and highly flexible, which means it can capture complex, non-linear relationships between multi-spectral Rrs inputs and chlorophyll without imposing a rigid functional form. At the same time, its Bayesian regularization – which shrinks the contribution of each individual tree – helps prevent overfitting even with a rich model structure. This balance between flexibility and regularization is well-suited to global ocean color data, where the true Rrs–Chl relationship is known to be non-linear and context-dependent, yet a model must generalize across diverse water types.

The interpretability of the BART model is illustrated by partial dependence analysis on the six input Rrs bands (centered approximately at 412, 443, 490, 510, 555, and 670 nm). The partial dependence plots reveal nuanced spectral–chlorophyll relationships and clear transitions in the dominant wavelength influence across different chlorophyll regimes. For example, the shorter blue wavelengths (412–443 nm) show a strong inverse influence on predicted $log(Chl_a)$ (higher blue reflectance corresponds to lower chlorophyll) that gradually saturates in the most oligotrophic waters, consistent with the leveling off of blue-to-green ratios at very low Chl. The green band (~555 nm) exhibits a more complex, non-monotonic effect: at low reflectance (clear water) its influence on the Chl estimate is minimal, but it becomes increasingly important through intermediate reflectance ranges – reflecting the band-difference signal exploited by the CI algorithm – and then diminishes for very high chlorophyll where green reflectance tends to flatten. In contrast, the red band (~670 nm) has virtually no effect on the model’s predictions until a threshold is reached at elevated chlorophyll concentrations, after which its influence steeply increases. This behavior aligns with optical expectations, since red reflectance is negligible in low-Chl waters but rises sharply once phytoplankton absorption in the blue saturates and biomass is high. These partial dependence results indicate that the BART model automatically learns the piecewise spectral logic that oceanographers often handle via separate algorithms (blue/green ratios for low-to-moderate Chl, red bands for high Chl). The ability to capture such regime-dependent responses in a single unified model, and to visualize them, underscores the interpretability and scientific insight provided by the BART approach.

Statement of Contribution
In this study, we develop and demonstrate a new global chlorophyll retrieval model based on Bayesian Additive Regression Trees implemented in PyMC. We train the BART model on a large, standardized dataset of satellite remote-sensing reflectance (Rrs) spectra matched with in situ chlorophyll measurements, using log₁₀-transformed Chl-a as the response to stabilize variance. The resulting model is applied globally to produce chlorophyll-a estimates from multi-spectral satellite data, with associated uncertainty estimates for each prediction. We show that this Bayesian tree-based model can serve as a general-purpose ocean color algorithm that is sensor-agnostic (provided reflectances are harmonized to common wavebands), interpretable, and uncertainty-aware. Unlike conventional empirical algorithms, the BART approach allows users to examine the inferred Rrs–Chl relationships and trust the model’s performance across regimes, while also quantifying confidence in each retrieval. This work thus contributes a novel methodological advance to satellite ocean color science: a unified chlorophyll retrieval model that marries the strengths of empirical algorithms (global applicability and simplicity) with the benefits of modern Bayesian machine learning (flexibility, interpretability, and rigorous uncertainty quantification). Our introduction of BART for global chlorophyll prediction opens the door for more robust monitoring of ocean biogeochemistry and improved integration of ocean color data into scientific and management applications.