### 2025-04-10 manuscript/manuscript.tex
finished introduction draft

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{agujournal2019}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation},
  pdfauthor={Erdem M. Karaköylü; Susanne E. Craig},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\journalname{Geophysical Research Letters}

\draftfalse

\begin{document}
\title{Shifting paradigms in Ocean Color: Bayesian Inference for
Uncertainty-Aware Chlorophyll Estimation}

\authors{Erdem M. Karaköylü\affil{1}, Susanne E. Craig\affil{2}}
\affiliation{1}{Consultant, }\affiliation{2}{NASA, }
\correspondingauthor{Erdem M. Karaköylü}{erdemk@protonmail.com}

\begin{abstract}
Placeholder
\end{abstract}

\section*{Plain Language Summary}
Placeholder

\subsection{Introduction}\label{introduction}

\subsubsection{Historical Context of Chlorophyll
Algorithms}\label{historical-context-of-chlorophyll-algorithms}

Satellite ocean color observations have long been fundamental for
monitoring marine ecosystems, as they enable the global estimation of
chlorophyll‑a---a key indicator of phytoplankton biomass and ocean
productivity. Early empirical algorithms, notably developed by O'Reilly
et al. (John E. O'Reilly et al., 1998; John E. O'Reilly et al., 2000),
established the \(OCx\) family (where \(x\) denotes the number of bands
used) of polynomial regression models. These models relate blue-to-green
reflectance ratios (after log‑transformation) to in situ chlorophyll‑a,
employing either straight band ratios (BR) or maximum band ratios
(MBR)---the latter selecting the highest available blue-to-green ratio
for any given observation as input to a high‑order polynomial. These
formulations have served as the operational foundation for chlorophyll‑a
products across a broad range of satellite ocean color sensors---from
the pioneering Coastal Zone Color Scanner (CZCS) through SeaWiFS, MODIS,
and MERIS to more recent missions---offering a straightforward and
robust approach for Case‑1 waters. However, their performance is more
limited in optically complex Case‑2 waters and remains sensitive to
atmospheric correction errors.

Subsequent refinements were introduced to address these deficiencies.
For example, Hu et al. (Hu et al., 2012) proposed a Color Index (CI)
formulation that employs a band‑difference approach to reduce
sensitivity to residual atmospheric errors and instrument noise, with
further improvements enhancing inter‑sensor agreement (Hu et al., 2019).
The increasing availability of calibration data (e.g.,
(\textbf{Valente2015?})) and ongoing algorithmic improvements have led
to the development of additional variants of the \(OCx\)
algorithms---specifically, the OC5 and OC6 formulations. O'Reilly and
Werdell (John E. O'Reilly \& Werdell, 2019) maintain that OC5 extends
the spectral basis by incorporating the 412\,nm band, thereby exploiting
its strong signal in clear, oligotrophic waters, while OC6 replaces the
traditional denominator with the mean of the 555 and 670\,nm
reflectances, improving the dynamic range at low chlorophyll
concentrations. In total, (John E. O'Reilly \& Werdell, 2019) propose 65
versions of BR/MBR \(OCx\) type algorithms for 25 sensors---on average,
two or more variants per sensor. With this arsenal, it is hoped,
researchers are better equipped to address the wide array of bio‑optical
environments encountered in global ocean color applications.

\subsubsection{Limitations of Existing
Approaches}\label{limitations-of-existing-approaches}

Regrettably, the development of traditional ocean color algorithms
relies on a fundamental statistical error. This error originates in the
conflation of empirical frequencies --- derived from a limited, noisy
set of observations --- with the theoretical relative frequencies that
would be observed over an infinite number of trials, thereby ignoring
the intrinsic uncertainty and variability inherent in finite sampling.
The fallout has and continues to be extremely detrimental to
observational science. Traditional, frequentist approaches, where the
variability in sample frequencies is treated as negligible, yield
overconfident inferences, and allegedgly significant findings that do
not hold water, in time leading to replication crises.

In statistical modeling, the goal is to assess how well a hypothesized
model \(M\) represents the relationship between independent and
dependent variables based on collected data \(D\). Frequentist methods
focus exclusively on the likelihood \(p(D|M)\), estimating model
parameters by maximizing this probability. This approach assumes that,
with a sufficiently large dataset, the chosen parameters reliably
reflect the true system. However, in practice, any finite dataset
carries inherent uncertainty and noise. As a consequence, models
developed under this framework tend to be overconfident - yielding point
estimates that do not adequately capture the variability of
\emph{in-situ} observations - and may not generalize well to new data.
Ad hoc attempts to generate uncertainty, such as bootstrapping, further
compound the problem by resampling the flawed, finite observations and
thus amplifying any existing biases. This overconfidence, coupled with
unverified assumptions presented as objectivity, can lead to misleading
inferences and contribute to replication challenges. The historical
process by which these assumptions have been enshrined is fascinating in
its own right, and we invite interested readers to consult works such as
McGrayne (\textbf{mcgrayne?}) and Clayton (\textbf{clayton?}).

As elaborated extensively by Jaynes (\textbf{jaynes2003probability?}),
and more recently by practitioners like Gelman
(\textbf{gelman2013bayesian?}), and McElreath
(\textbf{mcelreath2015statistical?}), probability is best understood as
a degree of knowledge rather than as a long-run frequency. Accordingly
in this Bayesian framework, the prior \(p(M)\) encodes our current state
of knowledge about the model's parameters before \(D\) is observed. This
knowledge can be elicited from experts, or if none exixts weakly
informative priors can be used. Note that whether \(M\) is based on
empirically observed relationships, analytical understanding, or
something in between is irrelevant to its integration into this
framework. Once the data \(D\) are collected, Bayes' theorem combines
the likelihood \(p(D|M)\) with the prior \(p(M)\) to yield the posterior
distribution \(p(M|D)\), which represents an updated level of knowledge
about the model's parameters. Crucially this posterior can then serve as
the new prior when new data become available, allowing for sequential
integration of new information. The by-product of this approach is
uncertainty estimates that more accurately reflect the complexities of
the underlying system as currently understood. In effect, rather than
relying on single point estimates based on \(p(D|M)\), the posterior
\(p(M|D)\) can be mined to retrieve a range of plausible parameter
values through credible intervals offering a more transparent and robust
basis for inference. This principled integration of prior knowledge with
observed data is central to our approach, as detailed further in the
Supplementary Material.

In this paper we recast a number of OCx-like algorithms in the Bayesian
framework. We also explore a different kind of

\subsubsection{Motivation for Bayesian Tree-Based
Models}\label{motivation-for-bayesian-tree-based-models}

To address these limitations, we explore Bayesian tree-based modeling --
specifically, Bayesian Additive Regression Trees (BART) -- as a novel
approach for ocean color chlorophyll retrieval. BART offers several
advantages for this application:

Direct Bayesian inference: As a Bayesian ensemble method, BART directly
targets the posterior distribution \(p(Chl | Rrs)\) rather than a purely
empirical fit. The model inherently provides a probabilistic mapping
from reflectance to chlorophyll. Rather than producing a single best-fit
value, BART returns a posterior mean and credible interval for each
prediction, naturally quantifying predictive uncertainty.

Uncertainty quantification: The BART framework includes uncertainty in
its predictions by design. Every estimated chlorophyll value comes with
a confidence range (e.g., 95\% credible interval) reflecting model
uncertainty and observational noise. This is a key improvement over
deterministic algorithms, enabling more reliable use of the data for
scientific and operational purposes where understanding uncertainty is
crucial e.g.~for forecasting trends.

Interpretability: Being a sum-of-trees model, BART can be interrogated
to understand the influence of each input variable. Tools such as
partial dependence plots allow us to visualize the relationship between
each spectral band and the predicted chlorophyll, marginalizing over
other bands. In contrast to high-order polynomial coefficients, these
tree-based partial dependencies are intuitively interpretable. For
instance, one can observe how changes in a particular waveband's
reflectance (holding others average) drive the chlorophyll estimate,
revealing any thresholds or non-linear responses.

BART's modeling approach is non-parametric and highly flexible, which
means it can capture complex, non-linear relationships between
multi-spectral Rrs inputs and chlorophyll without imposing a rigid
functional form. At the same time, its Bayesian regularization -- which
shrinks the contribution of each individual tree -- helps prevent
overfitting even with a rich model structure. This balance between
flexibility and regularization is well-suited to global ocean color
data, where the true Rrs--Chl relationship is known to be non-linear and
context-dependent, yet a model must generalize across diverse water
types.

The interpretability of the BART model is illustrated by partial
dependence analysis on the six input Rrs bands (centered approximately
at 412, 443, 490, 510, 555, and 670 nm). The partial dependence plots
reveal nuanced spectral--chlorophyll relationships and clear transitions
in the dominant wavelength influence across different chlorophyll
regimes. For example, the shorter blue wavelengths (412--443 nm) show a
strong inverse influence on predicted \(log(Chl_a)\) (higher blue
reflectance corresponds to lower chlorophyll) that gradually saturates
in the most oligotrophic waters, consistent with the leveling off of
blue-to-green ratios at very low Chl. The green band (\textasciitilde555
nm) exhibits a more complex, non-monotonic effect: at low reflectance
(clear water) its influence on the Chl estimate is minimal, but it
becomes increasingly important through intermediate reflectance ranges
-- reflecting the band-difference signal exploited by the CI algorithm
-- and then diminishes for very high chlorophyll where green reflectance
tends to flatten. In contrast, the red band (\textasciitilde670 nm) has
virtually no effect on the model's predictions until a threshold is
reached at elevated chlorophyll concentrations, after which its
influence steeply increases. This behavior aligns with optical
expectations, since red reflectance is negligible in low-Chl waters but
rises sharply once phytoplankton absorption in the blue saturates and
biomass is high. These partial dependence results indicate that the BART
model automatically learns the piecewise spectral logic that
oceanographers often handle via separate algorithms (blue/green ratios
for low-to-moderate Chl, red bands for high Chl). The ability to capture
such regime-dependent responses in a single unified model, and to
visualize them, underscores the interpretability and scientific insight
provided by the BART approach.

\section{Data Preprocessing}

Data for this study were acquired from multiple satellite ocean color
sensors and corresponding in situ chlorophyll-(a) measurements obtained
from sources such as the NASA Bio-Optical Marine Algorithm Data set
(NOMAD) and the compilation by Valente et al.~(2015). To ensure
consistency across sensors, the spectral reflectance data ((R\_\{rs\}))
were interpolated as needed to common wavelength centers.

For the empirical (OCx) formulation, blue-to-green band ratios were
computed for each observation. In particular, the maximum band ratio
(MBR) was determined by taking the highest value among the available
blue-band ratios (e.g., (Rrs(443)/Rrs(555)), (Rrs(490)/Rrs(555)), and
(Rrs(510)/Rrs(555))). This maximum value was then log-transformed: \[
\log R = \log_{10}\left(\frac{R_{rs}(\lambda_{\text{blue}})}{R_{rs}(555)}\right).
\]

For the Color Index (CI) formulation of Hu et al.~(2012), the CI was
calculated as: \[
\text{CI} = R_{rs}(555) - \left[\,R_{rs}(443) + \frac{555-443}{670-443}\Bigl(R_{rs}(670) - R_{rs}(443)\Bigr)\,\right],
\] and the corresponding in situ chlorophyll-(a) concentrations were
log-transformed: \[
\log \text{Chl} = \log_{10}(\text{Chl}).
\]

These transformations standardize the data to a common scale, ensuring
that variability is appropriately captured for subsequent regression and
uncertainty quantification. Detailed descriptions of the interpolation
methods and quality control procedures are provided in the Supplementary
Material.

\section{Statement of Contribution}\label{statement-of-contribution}

In this study, we develop and demonstrate a new global chlorophyll
retrieval model based on Bayesian Additive Regression Trees implemented
in PyMC. We train the BART model on a large, standardized dataset of
satellite remote-sensing reflectance (Rrs) spectra matched with in situ
chlorophyll measurements, using log₁₀-transformed Chl-a as the response
to stabilize variance. The resulting model is applied globally to
produce chlorophyll-a estimates from multi-spectral satellite data, with
associated uncertainty estimates for each prediction. We show that this
Bayesian tree-based model can serve as a general-purpose ocean color
algorithm that is sensor-agnostic (provided reflectances are harmonized
to common wavebands), interpretable, and uncertainty-aware. Unlike
conventional empirical algorithms, the BART approach allows users to
examine the inferred Rrs--Chl relationships and trust the model's
performance across regimes, while also quantifying confidence in each
retrieval. This work thus contributes a novel methodological advance to
satellite ocean color science: a unified chlorophyll retrieval model
that marries the strengths of empirical algorithms (global applicability
and simplicity) with the benefits of modern Bayesian machine learning
(flexibility, interpretability, and rigorous uncertainty
quantification). Our introduction of BART for global chlorophyll
prediction opens the door for more robust monitoring of ocean
biogeochemistry and improved integration of ocean color data into
scientific and management applications.

\section{Acknowledgments}\label{acknowledgments}

\section{Open research}\label{open-research}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\vspace{1em}

\bibitem[\citeproctext]{ref-hu2012novel}
Hu, C., Lee, Z., \& Franz, B. (2012). Chlorophyll aalgorithms for
oligotrophic oceans: A novel approach based on three-band reflectance
difference. \emph{Journal of Geophysical Research: Oceans},
\emph{117}(C1).
https://doi.org/\url{https://doi.org/10.1029/2011JC007395}

\bibitem[\citeproctext]{ref-hu2019improving}
Hu, C., Feng, L., Lee, Z., Franz, B. A., Bailey, S. W., Werdell, P. J.,
\& Proctor, C. W. (2019). Improving satellite global chlorophyll a data
products through algorithm refinement and data recovery. \emph{Journal
of Geophysical Research: Oceans}, \emph{124}(3), 1524--1543.
https://doi.org/\url{https://doi.org/10.1029/2019JC014941}

\bibitem[\citeproctext]{ref-oreilly2019}
O'Reilly, John E., \& Werdell, P. J. (2019). Chlorophyll algorithms for
ocean color sensors - OC4, OC5 \& OC6. \emph{Remote Sensing of
Environment}, \emph{229}, 32--47.
https://doi.org/\url{https://doi.org/10.1016/j.rse.2019.04.021}

\bibitem[\citeproctext]{ref-oreilly1998}
O'Reilly, John E., Maritorena, S., Mitchell, B. G., Siegel, D. A.,
Carder, K. L., Garver, S. A., et al. (1998). Ocean color chlorophyll
algorithms for SeaWiFS. \emph{Journal of Geophysical Research: Oceans},
\emph{103}(C11), 24937--24953.
https://doi.org/\url{https://doi.org/10.1029/98JC02160}

\bibitem[\citeproctext]{ref-oreilly2000}
O'Reilly, John E., Maritorena, S., Siegel, D. A., O'Brien, M. C., Toole,
D., Mitchell, B. G., et al. (2000). Ocean color chlorophyll a algorithms
for SeaWiFS, OC2, and OC4: Version 4. \emph{SeaWiFS Postlaunch
Calibration and Validation Analyses, Part}, \emph{3}, 9--23.

\end{CSLReferences}

\end{document}