### 2025-04-18 manuscript/manuscript.tex
Cleaned introduction

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{agujournal2019}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation},
  pdfauthor={Erdem M. Karaköylü; Susanne E. Craig},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\journalname{Geophysical Research Letters}

\draftfalse

\begin{document}
\title{Shifting paradigms in Ocean Color: Bayesian Inference for
Uncertainty-Aware Chlorophyll Estimation}

\authors{Erdem M. Karaköylü\affil{1}, Susanne E. Craig\affil{2}}
\affiliation{1}{Independant Consultant, }\affiliation{2}{NASA, }
\correspondingauthor{Erdem M. Karaköylü}{erdemk@protonmail.com}

\begin{abstract}
Placeholder
\end{abstract}

\section*{Plain Language Summary}
Placeholder

\section{Introduction}\label{introduction}

\subsection{Historical Context of Chlorophyll
Algorithms}\label{historical-context-of-chlorophyll-algorithms}

Satellite ocean color observations have long been fundamental for
monitoring marine ecosystems, as they enable the global estimation of
chlorophyll‑a (\(Chl_a\)) ---a key indicator of phytoplankton biomass
and ocean productivity. Early empirical algorithms, notably developed by
O'Reilly et al. (John E. O'Reilly et al., 1998; John E. O'Reilly et al.,
2000), established the \(OCx\) family (where \(x\) denotes the number of
bands used) of polynomial regression models. These models relate
blue-to-green reflectance ratios (after log‑transformation) to in situ
\(Chl_a\), employing either straight band ratios (BR) or maximum band
ratios (MBR)---the latter selecting the highest available blue-to-green
ratio for any given observation as input to a high‑order polynomial.
These formulations have served as the operational foundation for
chlorophyll‑a products across a broad range of satellite ocean color
sensors---from the pioneering Coastal Zone Color Scanner (CZCS) through
SeaWiFS, MODIS, and MERIS to more recent missions---offering a
straightforward and robust approach for Case‑1 waters. However, their
performance is more limited in optically complex Case‑2 waters and
remains sensitive to atmospheric correction errors.

Subsequent refinements were introduced to address these deficiencies.
For example, Hu et al. (Hu et al., 2012) proposed a Color Index (CI)
formulation that employs a band‑difference approach to reduce
sensitivity to residual atmospheric errors and instrument noise, with
further improvements enhancing inter‑sensor agreement (Hu et al., 2019).
The increasing availability of calibration data (e.g.,
(\textbf{Valente2015?})) and ongoing algorithmic improvements have led
to the development of additional variants of the \(OCx\)
algorithms---specifically, the OC5 and OC6 formulations. O'Reilly and
Werdell (John E. O'Reilly \& Werdell, 2019) maintain that OC5 extends
the spectral basis by incorporating the 412\,nm band, thereby exploiting
its strong signal in clear, oligotrophic waters, while OC6 replaces the
traditional denominator with the mean of the 555 and 670\,nm
reflectances, with the aim of improving the dynamic range at low
chlorophyll concentrations. In total, (John E. O'Reilly \& Werdell,
2019) propose 65 versions of BR/MBR \(OCx\) type algorithms for 25
sensors---on average, two or more variants per sensor. With this
arsenal, it is hoped, researchers are better equipped to address the
wide array of bio‑optical environments encountered in global ocean color
applications.

\subsection{Limitations of Existing
Approaches}\label{limitations-of-existing-approaches}

Regrettably, the development of traditional ocean color predictive
models relies on a fundamental statistical error that plagues most of
observational science today; that of conflating sampling probability,
with inferential probability(Scheemaekere \& Szafarz (2011).) Consider a
dataset \(D\) that consists of input-output pairs - e.g.~Remote sensing
reflectance (Rrs) and chlorophyll-\emph{a} concentration \(Chl_a\) - and
a model \(M\), such as OCx, hypothesized to describe the statistical
association between them. The sampling probability \(p(D|M)\) represents
the probability of observing the data \(D\) if model \(M\) were
``true''. Standard model fitting maximizes this quantity (i.e.~the
likelihood) by tuning the parameters of \(M\).

The unspoken assumption is that the model that best fits the data also
best represents the underlying process. This constitutes an epistemic
fallacy; substituting \(p(D|M)\) for \(p(M|D)\), in a violation of the
rules of conditional probability inversion as framed by Bayes' theorem.
While in well-behaved data-rich problems, the maxima of both expressions
may coincide, this is an exception, not the rule. The consequences of
this mistake are far-reaching as argued by (Clayton, 2022), with
implications extending from scientific modeling to medicine, law, and
public policy.

In science, this fallacy contributes to models fail to generalize well,
an impetus to produce ad-hoc or post-hoc uncertainty quantification, and
yield published results that resist replication(Cobey et al. (2024)).
These limiations are not confined to traditional hypothesis testing;
they are embedded in the training.; they are embedded in the training
and deployment of modern machine learning models as well. In regression
and classification tasks, likelihood maximization is treated as
sufficient for inference, even though it results in a single point
estimate and fails to account for parameter uncertainty or model
plausibility.

This epistemic shortcut has already been explicitly critiqued in the
machine learning community. Gal (Gal, 2016) and Ghahramani
(\textbf{gahramani2015probabilistic?}) emphasize that most ML models
discard uncertainty entirely and treat the output of optimization as
inference. This leads to overconfident predctions and brittle
generalization, echoing Clayton's concerncs. Bishop (Bishop, 2006) also
distiguishes between the utility of models and the inferential core
required to quantify uncertainty, underscoring that likelihood is not
enough and that the Bernoulli fallacy pervades the very logic of applied
machine learning.

\subsection{Overcoming limitations}\label{overcoming-limitations}

There have been attempts to address these issues. (Seegers et al., 2018)
have proposed alternative metrics to circumvent the inadequate
assumptions of the frequentist approach. Others have tried to go a step
futher incorporate Bayesian concepts. E.g. (Frouin \& Pelletier, 2013)
have proposed a Bayesian inversion scheme for atmospheric correction.
(\textbf{shi2015?}) proposed a probabilistic method to merge data from
different sensors. (Craig \& Karaköylü, 2019) have proposed a Bayesian
neural network (BNN) approach using Hamiltonian Monte Carlo sampling to
retrieve Inherent Optical Properties (IOP) from Top-of-the-atmosphere
(TOA) radiance. (\textbf{werther2022?}) used Monte-Carlo dropout to
approximate a deep BNN. (Erickson et al., 2023) have proposed using
conjugate Gaussian prior and likelihood to frame the GIOP as Bayesian
model to predict IOPs with uncertainty. (\textbf{hammout2024?}) have
proposed a BNN approximation via Stochastic Variational Inference to
predict \(Chl_a\) from ocean color observations. Yet most of these
approaches retain variable levels of frequentism by applying only part
of what is now commonly referred to as the Bayesian
workflow(\textbf{bgelman2019?}).

\subsection{In search of our Bayesian
workflow}\label{in-search-of-our-bayesian-workflow}

The Bayesian workflow is a recognizable and unifying way of approaching
statistical modeling, initially championed by statisticians at Columnbia
University (\textbf{gelman\_bda?}) subsequently quick endorsed and
further developed by researchers dealing with scant and noisy data (e.g.
(\textbf{mcelreath\_stat\_rethink?})). That is not to say that all steps
are applicable to all fields of research and all intents. The workflow
can and should be tailored to the researcher's field. (Wolkovich et al.,
2024) has proposed a four-step process for Ecological modeling. However
there are some core steps that should be followed. (1) Begin with one,
or better yet, more than one conceptual models and code them up. Model
building include prior formulation to encode existing knowledge. Priors
are part of the model structure and make the researcher's assumptions
transparent, providing a first avenue of critique and improvement. (2)
Check assumptions by a simulation process known as \emph{Prior
Predictive Checks}. Bayesian models are generative, meaning they can run
on empty, that is produce results without data. This allows a sanity
check of the built-in assumptions. Non-sensical simulation results
indicate assumptions must be revisited. (3) Collect data and conduct
exploratory data analysis. Knowing the data will be critical to
understanding potential problems during fitting; e..g multicollinearity
of input features. (4) Diagnostics; the fitting process and the
resulting posterior distribution provide rich constructs that can be
mined for a great deal of information. Most modern packages (e.g.~Stan,
PyMC, Numpyro, etc) offer a great many tools to extract insights. (5)
Model compasison and selection. Often a model will perform markedly
better, which is useful in terms predictive performance, but more
importantly in terms of generating insights. (6) Sequential update; a
model's output, the posterior distribution becomes the new prior every
time new data becomes available for a new sequence of model fitting.

In this paper we leverage past work and recast some of the pre-existing
\(OCx\) models into their probabilistic version. We also propose
alternative models, namely Bayesian Additive Regression Trees (BART,
(Chipman et al., 2010)) as a robust flexible method that has been used
for diverse application ((Linero, 2017).)

\section{Methods}\label{methods}

\subsection{Prior Elicitation and Model
Formulation}\label{prior-elicitation-and-model-formulation}

\subsection{Data Preprocessing}\label{data-preprocessing}

\subsection{}\label{section}

\section{Data Preprocessing}

Data for this study were acquired from multiple satellite ocean color
sensors and corresponding in situ chlorophyll-(a) measurements obtained
from sources such as the NASA Bio-Optical Marine Algorithm Data set
(NOMAD) and the compilation by Valente et al.~(2015). To ensure
consistency across sensors, the spectral reflectance data ((R\_\{rs\}))
were interpolated as needed to common wavelength centers.

For the empirical (OCx) formulation, blue-to-green band ratios were
computed for each observation. In particular, the maximum band ratio
(MBR) was determined by taking the highest value among the available
blue-band ratios (e.g., (Rrs(443)/Rrs(555)), (Rrs(490)/Rrs(555)), and
(Rrs(510)/Rrs(555))). This maximum value was then log-transformed: \[
\log R = \log_{10}\left(\frac{R_{rs}(\lambda_{\text{blue}})}{R_{rs}(555)}\right).
\]

For the Color Index (CI) formulation of Hu et al.~(2012), the CI was
calculated as: \[
\text{CI} = R_{rs}(555) - \left[\,R_{rs}(443) + \frac{555-443}{670-443}\Bigl(R_{rs}(670) - R_{rs}(443)\Bigr)\,\right],
\] and the corresponding in situ chlorophyll-(a) concentrations were
log-transformed: \[
\log \text{Chl} = \log_{10}(\text{Chl}).
\]

These transformations standardize the data to a common scale, ensuring
that variability is appropriately captured for subsequent regression and
uncertainty quantification. Detailed descriptions of the interpolation
methods and quality control procedures are provided in the Supplementary
Material.

\section{Statement of Contribution}\label{statement-of-contribution}

In this study, we develop and demonstrate a new global chlorophyll
retrieval model based on Bayesian Additive Regression Trees implemented
in PyMC. We train the BART model on a large, standardized dataset of
satellite remote-sensing reflectance (Rrs) spectra matched with in situ
chlorophyll measurements, using log₁₀-transformed Chl-a as the response
to stabilize variance. The resulting model is applied globally to
produce chlorophyll-a estimates from multi-spectral satellite data, with
associated uncertainty estimates for each prediction. We show that this
Bayesian tree-based model can serve as a general-purpose ocean color
algorithm that is sensor-agnostic (provided reflectances are harmonized
to common wavebands), interpretable, and uncertainty-aware. Unlike
conventional empirical algorithms, the BART approach allows users to
examine the inferred Rrs--Chl relationships and trust the model's
performance across regimes, while also quantifying confidence in each
retrieval. This work thus contributes a novel methodological advance to
satellite ocean color science: a unified chlorophyll retrieval model
that marries the strengths of empirical algorithms (global applicability
and simplicity) with the benefits of modern Bayesian machine learning
(flexibility, interpretability, and rigorous uncertainty
quantification). Our introduction of BART for global chlorophyll
prediction opens the door for more robust monitoring of ocean
biogeochemistry and improved integration of ocean color data into
scientific and management applications.

\section{Acknowledgments}\label{acknowledgments}

\section{Open research}\label{open-research}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\vspace{1em}

\bibitem[\citeproctext]{ref-baker2016}
Baker, M. (2016). 1,500 scientists lift the lid on reproducibility.
\emph{Nature}, \emph{533}(7604), 452--454.
\url{https://doi.org/10.1038/533452a}

\bibitem[\citeproctext]{ref-bishop2006pattern}
Bishop, C. M. (2006). \emph{Pattern recognition and machine learning}.
Springer.

\bibitem[\citeproctext]{ref-chipman2010bart}
Chipman, H. A., George, E. I., \& McCulloch, R. E. (2010). BART:
Bayesian additive regression trees. \emph{The Annals of Applied
Statistics}, \emph{4}(1), 266--298.
\url{https://doi.org/10.1214/09-AOAS285}

\bibitem[\citeproctext]{ref-clayton2022bernoulli}
Clayton, A. (2022). \emph{Bernoulli's fallacy: Statistical illogic and
the crisis of modern science}. Columbia University Press. Retrieved from
\url{https://books.google.com/books?id=BT4CzwEACAAJ}

\bibitem[\citeproctext]{ref-cobey2024biomedical}
Cobey, K. D., Ebrahimzadeh, S., Page, M. J., Thibault, R. T., Nguyen,
P.-Y., Abu-Dalfa, F., \& Moher, D. (2024). Biomedical researchers'
perspectives on the reproducibility of research. \emph{PLoS Biology},
\emph{22}(11), e3002870.

\bibitem[\citeproctext]{ref-craig2019}
Craig, S. E., \& Karaköylü, E. M. (2019). Bayesian models for deriving
biogeochemical information from satellite ocean color.
\emph{EarthArXiv}. \url{https://doi.org/10.31223/osf.io/shp6y}

\bibitem[\citeproctext]{ref-erickson2023}
Erickson, Z. K., McKinna, L. I. W., Werdell, P. J., \& Cetinić, I.
(2023). Bayesian approach to a generalized inherent optical property
model. \emph{Optics Express}, \emph{31}, 22790--22801.
\url{https://doi.org/10.1364/oe.486581}

\bibitem[\citeproctext]{ref-frouin2013}
Frouin, R., \& Pelletier, B. (2013). Bayesian methodology for ocean
color remote sensing. Retrieved from
\url{https://hal.archives-ouvertes.fr/hal-00822032}

\bibitem[\citeproctext]{ref-gal2016uncertainty}
Gal, Y. (2016). \emph{Uncertainty in deep learning} (PhD thesis).
University of Cambridge.

\bibitem[\citeproctext]{ref-hu2012novel}
Hu, C., Lee, Z., \& Franz, B. (2012). Chlorophyll aalgorithms for
oligotrophic oceans: A novel approach based on three-band reflectance
difference. \emph{Journal of Geophysical Research: Oceans},
\emph{117}(C1).
https://doi.org/\url{https://doi.org/10.1029/2011JC007395}

\bibitem[\citeproctext]{ref-hu2019improving}
Hu, C., Feng, L., Lee, Z., Franz, B. A., Bailey, S. W., Werdell, P. J.,
\& Proctor, C. W. (2019). Improving satellite global chlorophyll a data
products through algorithm refinement and data recovery. \emph{Journal
of Geophysical Research: Oceans}, \emph{124}(3), 1524--1543.
https://doi.org/\url{https://doi.org/10.1029/2019JC014941}

\bibitem[\citeproctext]{ref-jaynes2003probability}
Jaynes, E. T., \& Bretthorst, G. L. (2003). \emph{Probability theory:
The logic of science}. Cambridge University Press. Retrieved from
\url{https://books.google.com/books?id=tTN4HuUNXjgC}

\bibitem[\citeproctext]{ref-linero2017}
Linero, A. R. (2017). A review of tree-based bayesian methods.
\emph{Communications for Statistical Applications and Methods},
\emph{24}, 543--559. \url{https://doi.org/10.29220/CSAM.2017.24.6.543}

\bibitem[\citeproctext]{ref-oreilly2019}
O'Reilly, John E., \& Werdell, P. J. (2019). Chlorophyll algorithms for
ocean color sensors - OC4, OC5 \& OC6. \emph{Remote Sensing of
Environment}, \emph{229}, 32--47.
https://doi.org/\url{https://doi.org/10.1016/j.rse.2019.04.021}

\bibitem[\citeproctext]{ref-oreilly1998}
O'Reilly, John E., Maritorena, S., Mitchell, B. G., Siegel, D. A.,
Carder, K. L., Garver, S. A., et al. (1998). Ocean color chlorophyll
algorithms for SeaWiFS. \emph{Journal of Geophysical Research: Oceans},
\emph{103}(C11), 24937--24953.
https://doi.org/\url{https://doi.org/10.1029/98JC02160}

\bibitem[\citeproctext]{ref-oreilly2000}
O'Reilly, John E., Maritorena, S., Siegel, D. A., O'Brien, M. C., Toole,
D., Mitchell, B. G., et al. (2000). Ocean color chlorophyll a algorithms
for SeaWiFS, OC2, and OC4: Version 4. \emph{SeaWiFS Postlaunch
Calibration and Validation Analyses, Part}, \emph{3}, 9--23.

\bibitem[\citeproctext]{ref-DeScheemaekere2011}
Scheemaekere, X. D., \& Szafarz, A. (2011). The inference fallacy from
bernoulli to kolmogorov. Retrieved from
\url{https://www.researchgate.net/publication/254450993_The_Inference_Fallacy_From_Bernoulli_to_Kolmogorov}

\bibitem[\citeproctext]{ref-seegers2018}
Seegers, B. N., Stumpf, R. P., Schaeffer, B. A., Loftin, K. A., \&
Werdell, P. J. (2018). Performance metrics for the assessment of
satellite data products: An ocean color case study. \emph{Opt. Express},
\emph{26}(6), 7404--7422. \url{https://doi.org/10.1364/OE.26.007404}

\bibitem[\citeproctext]{ref-wolkovich2024fourstepbayesianworkflowimproving}
Wolkovich, E., Davies, T. J., Pearse, W. D., \& Betancourt, M. (2024). A
four-step bayesian workflow for improving ecological science. Retrieved
from \url{https://arxiv.org/abs/2408.02603}

\end{CSLReferences}

\end{document}