### 2025-05-16 manuscript/manuscript.qmd
updated manuscript

---
title: "Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation "
format:
    agu-pdf:
        keep-tex: true
    agu-html: default
author:  
  - name: Erdem M. Karaköylü
    orcid: 0000-0002-6156-1720
    corresponding: true
    email: erdemk@protonmail.com
    roles:
      - Investigator
      - Software
      - Visualization
      - Project administration
    affiliations:
      - Independant Researcher
  - name: Susanne E. Craig
    orcid: 0000-0002-8963-0951
    corresponding: false
    email: susanne.e.craig@nasa.gov
    roles:
      - Co-Investigator
    affiliations:
      - NASA/UMBC
abstract: Placeholder
plain-language-summary: Placeholder
keywords: ['Bayesian modeling', 'Ocean Color', ]
key-points:
  - Bayesian models have not been fully adopted in Ocean Color remote sensing yet.
  - A Bayesian workflow is a principled guide to a fully Bayesian study
  - 
bibliography: references.bib
citation:
  container-title: Geophysical Research Letters
keep-tex: true
number-sections: false
date: last-modified
---

## Introduction

Satellite ocean color remote sensing has long served as a cornerstone of marine ecosystem monitoring, offering global and synoptic coverage of surface ocean properties. Among these, chlorophyll-a ($Chl_a$) concentration remains a central metric, widely used as a proxy for phytoplankton biomass, primary production, and water quality. The retrieval of $Chl_a$ from ocean color data has evolved over decades, resulting in a diverse lineage of empirical and semi-empirical algorithms. The following section summarizes this historical development, which sets the stage for a critical examination of the statistical foundations underlying current approaches.

### Historical Context of Chlorophyll Algorithms

Satellite ocean color observations have long been fundamental for monitoring marine ecosystems, as they enable global estimation of chlorophyll‑a ($Chl_a$) — a key indicator of phytoplankton biomass and ocean productivity. Early empirical algorithms, notably developed by O’Reilly et al. [@oreilly1998; @oreilly2000], established the $OCx$ family (where $x$ denotes the number of bands used) of polynomial regression models. These models relate blue-to-green reflectance ratios (after log‑transformation) to in situ $Chl_a$, employing either straight band ratios (BR) or maximum band ratios (MBR)—the latter selecting the highest available blue-to-green ratio for any given observation as input to a high‑order polynomial. These formulations have served as the operational foundation for chlorophyll‑a products across a broad range of satellite ocean color sensors—from the pioneering Coastal Zone Color Scanner (CZCS) through SeaWiFS, MODIS, and MERIS to more recent missions—offering a straightforward and robust approach for Case‑1 waters. However, their performance is more limited in optically complex Case‑2 waters and remains sensitive to atmospheric correction errors.

Subsequent refinements were introduced to address these deficiencies. For example, Hu et al. [@hu2012novel] proposed a Color Index (CI) formulation that employs a band‑difference approach to reduce sensitivity to residual atmospheric errors and instrument noise, with further improvements enhancing inter‑sensor agreement [@hu2019improving]. The increasing availability of calibration data (e.g., [@Valente2015]) and ongoing algorithmic improvements have led to the development of additional variants of the $OCx$ algorithms—specifically, the OC5 and OC6 formulations. O’Reilly and Werdell [@oreilly2019] maintain that OC5 extends the spectral basis by incorporating the 412 nm band, thereby exploiting its strong signal in clear, oligotrophic waters, while OC6 replaces the traditional denominator with the mean of the 555 and 670 nm reflectances, with the aim of improving the dynamic range at low chlorophyll concentrations. In total, [@oreilly2019] propose 65 versions of BR/MBR $OCx$ type algorithms for 25 sensors—on average, two or more variants per sensor. With this arsenal, it is hoped, researchers are better equipped to address the wide array of bio‑optical environments encountered in global ocean color applications.

### Limitations of Existing Approaches

Regrettably, the development of traditional ocean color algorithms is grounded in a fundamental statistical error—one that pervades much of observational science: the conflation of sampling probability with inferential probability [@jaynes2003probability; @DeScheemaekere2011].

Consider a dataset $D$ composed of input–output pairs—e.g., remote sensing reflectance (Rrs) and chlorophyll-a concentration ($Chl_a$)—and a model $M$, such as OCx, posited to represent the relationship between them. The sampling probability $p(D \mid M)$ denotes the probability of observing data $D$ under the assumption that model $M$ is true. In standard model fitting, this likelihood is maximized by adjusting the parameters of $M$ to best explain the observed data.

This approach tacitly assumes that the model which best fits the data also most accurately represents the underlying generative process. This constitutes an epistemic fallacy—treating $p(D \mid M)$ as if it were $p(M \mid D)$—a direct violation of Bayes’ theorem and the rules governing conditional probability.

Although in well-behaved, data-rich cases—where the likelihood is regular, the signal strong, and the model adequately constrained—the maxima of $p(D \mid M)$ and $p(M \mid D)$ may coincide, this remains the exception—not the rule.

This mistake lies at the heart of what Clayton [@clayton2022bernoulli] terms the Bernoulli Fallacy: the widespread misinterpretation of likelihood as inference, or of data-fit as belief. As Clayton argues, this logical misstep has far-reaching consequences, with implications that extend beyond science to domains such as medicine, law, and public policy.

In scientific modeling, this fallacy contributes to poor generalization, drives the use of ad hoc or retrospective uncertainty quantification, and underlies many published results that later prove difficult to replicate [@baker2016; @cobey2024biomedical]. These limitations are not restricted to classical hypothesis testing; they persist in the training and deployment of modern machine learning models as well.

In regression and classification, maximizing likelihood is often treated as sufficient for inference—despite yielding only a single point estimate and ignoring both parameter uncertainty and the plausibility of alternative models.

This epistemic shortcut has been directly critiqued in the machine learning literature. Gal [@gal2016uncertainty] and Ghahramani [@gahramani2015probabilistic] point out that most ML models discard uncertainty altogether, treating the outcome of an optimization as if it were an inference. The result is overconfident predictions and brittle generalization—concerns that echo Clayton’s critique.

Bishop [@bishop2006pattern] similarly distinguishes between the utility of predictive models and the inferential scaffolding required to quantify uncertainty, reinforcing the notion that likelihood alone is insufficient—and that the Bernoulli Fallacy permeates much of applied machine learning.

### Overcoming limitations

There have been attempts to address these issues. [@seegers2018] have proposed alternative metrics to circumvent the inadequate assumptions of the frequentist approach. Others have tried to go a step futher incorporate Bayesian concepts. E.g. [ @frouin2013] have proposed a Bayesian inversion scheme for atmospheric correction. [@shi2015] proposed a probabilistic method to merge data from different sensors. [@craig2019] have proposed a Bayesian neural network (BNN) approach using Hamiltonian Monte Carlo sampling to retrieve Inherent Optical Properties (IOP) from Top-of-the-atmosphere (TOA) radiance. [@werther2022] used Monte-Carlo dropout to approximate a deep BNN. [@erickson2023] have proposed using conjugate Gaussian prior and likelihood to frame the GIOP as Bayesian model to predict IOPs with uncertainty. [@hammout2024] have proposed a BNN approximation via Stochastic Variational Inference to predict $Chl_a$ from ocean color observations. Yet most of these approaches retain variable levels of frequentism by applying only part of what is now commonly referred to as the Bayesian workflow[@bgelman2019].

### In search of our Bayesian workflow
The Bayesian workflow is a recognizable and unifying way of approaching statistical modeling, initially championed by statisticians at Columnbia University [@gelman_bda] subsequently quick endorsed and further developed by researchers dealing with scant and noisy data (e.g. [@mcelreath_stat_rethink]). That is not to say that all steps are applicable to all fields of research and all intents. The workflow can and should be tailored to the researcher's field. [@wolkovich2024fourstepbayesianworkflowimproving] has proposed a four-step process for Ecological modeling. 
However there are some core steps that should be followed. (1) Begin with one, or better yet,  more than one conceptual models and code them up.  Model building include prior formulation to encode existing knowledge. Priors are part of the model structure and make the researcher's assumptions transparent, providing a first avenue of critique and improvement. (2) Check assumptions by a simulation process known as *Prior Predictive Checks*. Bayesian models are generative, meaning they can run on empty, that is produce results without data. This allows a sanity check of the built-in assumptions. Non-sensical simulation results indicate assumptions must be revisited. (3) Collect data and conduct exploratory data analysis. Knowing the data will be critical to understanding potential problems during fitting; e..g multicollinearity of input features. (4) Diagnostics; the fitting process and the resulting posterior distribution provide rich  constructs that can be mined for a great deal of information. Most modern packages (e.g. Stan, PyMC, Numpyro, etc) offer a great many tools to extract insights. (5) Model compasison and selection. Often a model will perform markedly better, which is useful in terms predictive performance, but more importantly in terms of generating insights. (6) Sequential update; a model's output, the posterior distribution becomes the new prior every time new data becomes available for a new sequence of model fitting.

In this paper we leverage past work but also propose new approaches for predicting chlorophyll from sea surface reflectance (Rrs). Overall we propose three types of models.

## 2. Materials and Methods

This section outlines the study region and dataset used for model development and validation, the structure and rationale for the Bayesian models, and the approach used to compare their predictive performance. We emphasize a subset of four representative models in the main text, while full specifications and results for all seven models are provided in the Supplementary Material.

### 2.1 Study Area and Dataset

We used satellite and in situ data from the NOMAD v2 (2008) dataset to train and validate chlorophyll-a prediction models based on remote sensing reflectance (Rrs). Detailed preprocessing and matching procedures were applied to ensure high-quality paired observations suitable for statistical modeling.

<!-- To expand: summarize geographic scope if applicable, temporal resolution, types of chlorophyll measurements, sensor bands used -->

- NOMAD v2 dataset: in situ measurements and coincident satellite Rrs
- Removal of invalid or zero Rrs values
- Computation of Max Band Ratio (MBR)
- log-transformation of chlorophyll and MBR
- Creation of group labels based on spectral dominance
- Measurement method labels (fluorescence vs. HPLC)

### 2.2 Bayesian Modeling Framework

We adopt a Bayesian approach to flexibly capture the relationship between log-transformed MBR and log-transformed chlorophyll-a. Bayesian inference allows us to quantify uncertainty, incorporate prior knowledge, and rigorously compare alternative model structures using probabilistic criteria.

<!-- To expand: cite PyMC, mention benefits of full posterior inference, prior predictive checking, model flexibility -->

- Inference implemented using PyMC v5
- Models specified using a Truncated Normal likelihood
- Noise term (σ) modeled with increasing complexity
- Priors: weakly informative Normal for regression terms; Gamma for σ

### 2.3 Overview of Candidate Models

We developed seven Bayesian models of increasing complexity. These models vary in terms of functional form (polynomial vs. linear), structure (flat vs. hierarchical), error formulation (constant vs. heteroscedastic), and treatment of measurement-specific noise.

For brevity and clarity, we present four representative models in the main text:
- **Model 1**: Polynomial regression (Bayesian OC6)
- **Model 2**: Hierarchical linear regression grouped by Rrs band
- **Model 6**: Hierarchical linear regression with group-wise heteroscedasticity
- **Model 7**: Model 6 + additional noise term for fluorescence-based Chl

A summary table of all models, including the three omitted from the main text, is provided below. Full mathematical specifications, trace diagnostics, and model performance comparisons are included in the Supplement.

<!-- To expand: add table summarizing models (model ID, functional form, grouping, σ structure, measurement adjustment) -->

### 2.4 Representative Model Structures

#### 2.4.1 Model 1: Bayesian Polynomial OC6 Regression

This model mirrors the classical OC6 algorithm with a polynomial regression on log(MBR). It provides a simple, non-hierarchical baseline for model comparisons.

- Polynomial structure similar to O'Reilly et al. (2019)
- Constant σ
- No group structure or measurement adjustment

#### 2.4.2 Model 2: Hierarchical Linear Regression (4-band Grouping)

This model introduces partial pooling by grouping observations according to the dominant Rrs band used in the MBR numerator. It aims to address known discontinuities in empirical band-ratio algorithms.

- Partial pooling on slope and intercept terms
- Shared hyperpriors across groups
- Constant σ

#### 2.4.3 Model 6: Hierarchical Linear Regression with Group-wise Linear σ

This model extends Model 2 by allowing the noise term σ to vary by group as a log-linear function of MBR, capturing heteroscedasticity commonly observed in chlorophyll datasets.

- Group-specific linear model for log(σ)
- Partial pooling of σ slope/intercept parameters

#### 2.4.4 Model 7: Hierarchical Linear σ + Measurement-Method-Specific Noise

Building on Model 6, this model introduces an additional noise term for chlorophyll values derived from fluorescence, accounting for the increased uncertainty relative to HPLC-based measurements.

- Same structure as Model 6
- Additional variance term for fluorescence-flagged observations

### 2.5 Prior Specification and Inference Settings

All models use weakly informative priors and were fit using Hamiltonian Monte Carlo (NUTS) in PyMC. Convergence and effective sampling were confirmed via standard diagnostics.

<!-- To expand: mention specific prior distributions used, tuning parameters, number of chains/draws -->

- Priors: Normal for regression terms, Gamma for σ
- Chains, tuning steps, and draws per model
- Convergence assessed using R-hat and ESS
- Trace plots and diagnostics in Supplementary Material

### 2.6 Model Comparison and Predictive Evaluation

Models were compared using Leave-One-Out Cross Validation (LOO-CV) based on Pareto Smoothed Importance Sampling (PSIS), a robust criterion for evaluating out-of-sample predictive performance.

<!-- To expand: cite ArviZ, explain ELPD, SE(ΔELPD), ranking strategy -->

- Model comparison was performed using **Pareto Smoothed Importance Sampling Leave-One-Out Cross-Validation (PSIS-LOO-CV)**, implemented via the `arviz.loo()` function.
- Models were ranked based on their **Expected Log Predictive Density (ELPD)**, a Bayesian measure of out-of-sample predictive accuracy.
- Differences in ELPD between models were evaluated alongside their **standard errors (SE)** to assess whether improvements were credibly distinguishable.
- **Predictive accuracy** was assessed using:
  - **Posterior Predictive Checks (PPCs)**: to visualize the degree to which simulated data resembled observed outcomes.
  - **Highest Density Interval (HDI) coverage**: to evaluate the calibration of predictive intervals relative to held-out data.
- No frequentist criteria (e.g., RMSE, R²) were used at any stage of model evaluation or selection.

## 3. Results

This section presents convergence diagnostics, posterior inference, and model comparison using fully Bayesian metrics. No frequentist evaluation criteria were used. Instead, we rely on information-theoretic model comparison (LOO-CV), posterior predictive checks, and visualization of uncertainty via highest density intervals (HDIs).

### 3.1 MCMC Convergence

All four models converged successfully. The Gelman-Rubin statistic (R-hat ≈ 1.0) and large effective sample sizes (ESS > 1000) across all parameters indicate robust posterior estimation. Diagnostic plots and sampling traces are provided in the Supplementary Material.

- All R-hat < 1.01
- No divergences
- Sufficient tail and bulk ESS
- Full diagnostics in Supplement

### 3.2 Predictive Performance via Leave-One-Out Cross Validation

We compared models using Pareto Smoothed Importance Sampling Leave-One-Out Cross-Validation (PSIS-LOO-CV), a fully Bayesian approach to evaluating out-of-sample predictive accuracy.

**Table 1.** Leave-One-Out Comparison (ELPD ± SE):

<!-- Insert Quarto table of ELPD, ΔELPD, SE, p_loo -->

- Model 7 ranked highest in expected log predictive density (ELPD)
- Model 6 performed comparably, with ΔELPD within 1 SE
- Model 1 (OC6 polynomial) had the weakest predictive performance
- All models had acceptable Pareto-k diagnostics (see Supplement)

### 3.3 Posterior Predictive Checks (PPC)

We used posterior predictive checks (PPCs) to evaluate the ability of each model to reproduce key features of the observed log-chlorophyll distribution. These include:

- Empirical CDF comparisons
- HDI envelopes overlaid on the data
- Predictive median vs. observed values

**Figure 1.** Posterior predictive envelopes vs. observed log(Chl)  
**Figure 2.** ECDF overlays with posterior predictive means and HDIs  
**Figure 3.** Rug plot of observations with overlaid posterior predictive draws

Findings:
- Model 7 captures full data distribution, especially for fluorescence-labeled values
- Model 1 underestimates tails (under-dispersed)
- Model 6 accurately reflects group-level variance

### 3.4 Posterior Distributions and Interpretability

Posterior distributions for key parameters illustrate how model structure affects inference. We show both:
- Group-level slopes and intercepts (hierarchical structure)
- Group-wise σ (heteroscedasticity)
- Fluorescence-specific variance term (Model 7)

**Figure 4.** Forest plot of slope and intercept posteriors  
**Figure 5.** Posterior of σ across MBR groups  
**Figure 6.** Posterior of added noise term for fluorescence

Key insights:
- Rrs510-dominant group had the steepest slope (Model 2 & 6)
- Fluorescence noise term is credibly nonzero (Model 7)
- Posterior variance structure supports heteroscedasticity as a key feature

### 3.5 Summary of Bayesian Comparison

- **Model 7**: Best predictive performance (highest ELPD), best uncertainty calibration
- **Model 6**: Nearly equivalent performance; simpler if fluorescence uncertainty not needed
- **Model 2**: Gains from partial pooling; underperforms without variance modeling
- **Model 1**: Legacy structure underfits tails, overconfident predictions

All results support the Bayesian hierarchical modeling framework with heteroscedasticity and measurement-aware noise as optimal for chlorophyll-a estimation.

## 4. Discussion

This section interprets the results in the context of previous chlorophyll-a modeling work, with particular emphasis on the added value of Bayesian modeling. We highlight the importance of accounting for heteroscedasticity and measurement-type uncertainty, and discuss how this contributes to improved predictive reliability and scientific interpretability.

### 4.1 Value of Bayesian Framework over Classical Approaches

Our results demonstrate the clear advantages of fully Bayesian modeling for chlorophyll-a prediction from satellite Rrs. Unlike deterministic or frequentist models:

- Bayesian methods explicitly model parameter uncertainty and provide HDIs instead of misleading confidence intervals
- Posterior predictive distributions enable direct evaluation of model calibration
- Bayesian model comparison via LOO-CV avoids overfitting and guards against inappropriate complexity

<!-- Expand: reinforce critique of RMSE-based selection; cite key Bayesian modeling references -->

### 4.2 Importance of Hierarchical Structuring

Models that leveraged group-wise structure (based on dominant Rrs bands) consistently outperformed simpler counterparts.

- Hierarchical partial pooling improved both prediction and parameter identifiability
- Group-level trends reflected known biogeophysical distinctions (e.g., clear vs. turbid water types)
- Intercepts and slopes varied systematically across spectral regimes

<!-- Expand: relate group-level results to known Case 1 vs. Case 2 water differences -->

### 4.3 Role of Heteroscedasticity in Chlorophyll Retrieval

Accounting for non-constant noise was essential for accurate uncertainty quantification.

- Models with linear or hierarchical σ terms captured increased variance at higher MBR or Chl
- Constant-σ models (e.g., OC6) underrepresented predictive uncertainty, especially in extremes
- Explicit modeling of σ structure produced calibrated HDIs and superior posterior predictive checks

<!-- Expand: link back to observed data behavior and prior studies showing variance scaling with concentration -->

### 4.4 Measurement-Specific Error Modeling

Model 7, which introduced a separate noise term for fluorescence-derived Chl, outperformed all others.

- Fluorescence-based measurements have higher and more variable error
- This heterogeneity cannot be ignored in merged datasets
- Bayesian modeling allows this variance to be encoded explicitly without discarding data

<!-- Expand: cite studies on fluorescence bias, show that Bayesian methods allow uncertainty to “absorb” method bias -->

### 4.5 Implications for Satellite Algorithm Development

The findings support incorporating Bayesian elements into future operational chlorophyll algorithms.

- Hierarchical modeling offers a path to generalize across bioregions and sensors
- Posterior uncertainty estimates can support downstream applications (e.g., forecasting, ecological thresholds)
- Bayesian frameworks are sensor-agnostic: Rrs inputs can vary as long as priors are adapted

<!-- Optional: comment on feasibility of deploying these methods, or emulating them with approximate Bayesian techniques -->

### 4.6 Limitations and Future Work

As with any modeling effort, our study has limitations:

- All results are conditional on the data from NOMAD; further validation on independent datasets is needed
- More sophisticated noise structures (e.g., nonparametric σ) could be explored
- Spatial or spatio-temporal extensions were not considered, but would be feasible in PyMC

<!-- Expand: suggest extensions (Gaussian Processes, spatial priors, stacking), and invite reproducibility -->

## Acknowledgments

## Open research

## References {.unnumbered}

:::{#refs}

:::