### 2025-05-17 manuscript/manuscript.qmd
updated introduction

---
title: "Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation "
format:
    agu-pdf:
        keep-tex: true
    agu-html: default
author:  
  - name: Erdem M. Karaköylü
    orcid: 0000-0002-6156-1720
    corresponding: true
    email: erdemk@protonmail.com
    roles:
      - Investigator
      - Software
      - Visualization
      - Project administration
    affiliations:
      - Independant Researcher
  - name: Susanne E. Craig
    orcid: 0000-0002-8963-0951
    corresponding: false
    email: susanne.e.craig@nasa.gov
    roles:
      - Co-Investigator
    affiliations:
      - NASA/UMBC
abstract: Placeholder
plain-language-summary: Placeholder
keywords: ['Bayesian modeling', 'Ocean Color', ]
key-points:
  - Bayesian models have not been fully adopted in Ocean Color remote sensing yet.
  - A Bayesian workflow is a principled guide to a fully Bayesian study
  - 
bibliography: references.bib
citation:
  container-title: Geophysical Research Letters
keep-tex: true
number-sections: false
date: last-modified
---

## Introduction

Satellite ocean color remote sensing has long served as a cornerstone of marine ecosystem monitoring, offering global and synoptic coverage of surface ocean properties. Among these, chlorophyll-a ($Chl_a$) concentration remains a central metric, widely used as a proxy for phytoplankton biomass, primary production, and water quality. The retrieval of $Chl_a$ from ocean color data has evolved over decades, resulting in a diverse lineage of empirical and semi-empirical algorithms. The following section summarizes this historical development, which sets the stage for a critical examination of the statistical foundations underlying current approaches.

### Historical Context of Chlorophyll Algorithms

Satellite ocean color observations have long been fundamental for monitoring marine ecosystems, as they enable global estimation of chlorophyll‑a ($Chl_a$) — a key indicator of phytoplankton biomass and ocean productivity. Early empirical algorithms, notably developed by O’Reilly et al. [@oreilly1998; @oreilly2000], established the $OCx$ family (where $x$ denotes the number of bands used) of polynomial regression models. These models relate blue-to-green reflectance ratios (after log‑transformation) to in situ $Chl_a$, employing either straight band ratios (BR) or maximum band ratios (MBR)—the latter selecting the highest available blue-to-green ratio for any given observation as input to a high‑order polynomial. These formulations have served as the operational foundation for chlorophyll‑a products across a broad range of satellite ocean color sensors—from the pioneering Coastal Zone Color Scanner (CZCS) through SeaWiFS, MODIS, and MERIS to more recent missions—offering a straightforward and robust approach for Case‑1 waters. However, their performance is more limited in optically complex Case‑2 waters and remains sensitive to atmospheric correction errors.

Subsequent refinements were introduced to address these deficiencies. For example, Hu et al. [@hu2012novel] proposed a Color Index (CI) formulation that employs a band‑difference approach to reduce sensitivity to residual atmospheric errors and instrument noise, with further improvements enhancing inter‑sensor agreement [@hu2019improving]. The increasing availability of calibration data (e.g., [@Valente2015]) and ongoing algorithmic improvements have led to the development of additional variants of the $OCx$ algorithms—specifically, the OC5 and OC6 formulations. O’Reilly and Werdell [@oreilly2019] maintain that OC5 extends the spectral basis by incorporating the 412 nm band, thereby exploiting its strong signal in clear, oligotrophic waters, while OC6 replaces the traditional denominator with the mean of the 555 and 670 nm reflectances, with the aim of improving the dynamic range at low chlorophyll concentrations. In total, [@oreilly2019] propose 65 versions of BR/MBR $OCx$ type algorithms for 25 sensors—on average, two or more variants per sensor. With this arsenal, it is hoped, researchers are better equipped to address the wide array of bio‑optical environments encountered in global ocean color applications.

### Limitations of Existing Approaches

Regrettably, the development of traditional ocean color algorithms is grounded in a fundamental statistical error—one that pervades much of observational science: the conflation of sampling probability with inferential probability [@jaynes2003probability; @DeScheemaekere2011].

Consider a dataset $D$ composed of input–output pairs—e.g., remote sensing reflectance (Rrs) and chlorophyll-a concentration ($Chl_a$)—and a model $M$, such as OCx, posited to represent the relationship between them. The sampling probability $p(D \mid M)$ denotes the probability of observing data $D$ under the assumption that model $M$ is true. In standard model fitting, this likelihood is maximized by adjusting the parameters of $M$ to best explain the observed data.

This approach tacitly assumes that the model which best fits the data also most accurately represents the underlying generative process. This constitutes an epistemic fallacy—treating $p(D \mid M)$ as if it were $p(M \mid D)$—a direct violation of Bayes’ theorem and the rules governing conditional probability.

Although in well-behaved, data-rich cases—where the likelihood is regular, the signal strong, and the model adequately constrained—the maxima of $p(D \mid M)$ and $p(M \mid D)$ may coincide, this remains the exception—not the rule.

This mistake lies at the heart of what Clayton [@clayton2022bernoulli] terms the Bernoulli Fallacy: the widespread misinterpretation of likelihood as inference, or of data-fit as belief. As Clayton argues, this logical misstep has far-reaching consequences, with implications that extend beyond science to domains such as medicine, law, and public policy.

In scientific modeling, this fallacy contributes to poor generalization, drives the use of ad hoc or retrospective uncertainty quantification, and underlies many published results that later prove difficult to replicate [@baker2016; @cobey2024biomedical]. These limitations are not restricted to classical hypothesis testing; they persist in the training and deployment of modern machine learning models as well.

In regression and classification, maximizing likelihood is often treated as sufficient for inference—despite yielding only a single point estimate and ignoring both parameter uncertainty and the plausibility of alternative models.

This epistemic shortcut has been directly critiqued in the machine learning literature. Gal [@gal2016uncertainty] and Ghahramani [@gahramani2015probabilistic] point out that most ML models discard uncertainty altogether, treating the outcome of an optimization as if it were an inference. The result is overconfident predictions and brittle generalization—concerns that echo Clayton’s critique.

Bishop [@bishop2006pattern] similarly distinguishes between the utility of predictive models and the inferential scaffolding required to quantify uncertainty, reinforcing the notion that likelihood alone is insufficient—and that the Bernoulli Fallacy permeates much of applied machine learning.

### Overcoming limitations

Several recent efforts have attempted to address the limitations of classical retrieval models. For instance, [@seegers2018] proposed alternative evaluation metrics to move beyond restrictive frequentist assumptions. Others have incorporated Bayesian elements into the modeling pipeline: [@frouin2013] applied Bayesian inversion for atmospheric correction, [@shi2015] used probabilistic fusion for multi-sensor data, and [@craig2019] employed Hamiltonian Monte Carlo to train Bayesian neural networks (BNNs) for retrieving inherent optical properties (IOPs) from top-of-atmosphere radiance. Similarly, [@werther2022] introduced Monte Carlo dropout as an approximate BNN strategy, while [@erickson2023] recast the Generalized Inherent Optical Property (GIOP) framework using conjugate Bayesian linear models. Most recently, [@hammout2024] developed a BNN surrogate using stochastic variational inference for chlorophyll-a prediction from Rrs.

While these studies mark important progress, they often adopt only isolated components of what has come to be known as the Bayesian workflow [@bgelman2019]. Key aspects—such as principled prior specification, posterior predictive checking, and formal model comparison—remain largely absent. As a result, uncertainty is frequently approximated rather than inferred, and model structure is rarely treated as a variable to be interrogated.

This paper builds on prior work and extends it by applying a fully Bayesian modeling framework for chlorophyll-a prediction from sea surface reflectance (Rrs), with explicit treatment of uncertainty, model complexity, and measurement error.

## 2. Materials and Methods

This section outlines the dataset, model structures, inference methods, and model evaluation criteria used in this study. I focus on a fully Bayesian approach to modeling chlorophyll-a from satellite remote sensing reflectance (Rrs), emphasizing transparency, uncertainty quantification, and principled model comparison. Four representative models are presented in the main text, while all seven model formulations and results are included in the Supplement.

### 2.1 Study Area and Dataset

I used the NOMAD v2 (2008) dataset, which includes quality-controlled in situ measurements of chlorophyll-a matched with coincident satellite Rrs data. These measurements span a range of oceanographic conditions, enabling the development of models that generalize across water types.

To prepare the data:

- **Chlorophyll-a**:  
  - Measurements obtained using both HPLC and fluorescence methods  
  - Combined into a single log-transformed target variable (log₁₀ Chl)  
  - A binary flag identifies the measurement method for each sample

- **Remote sensing reflectance (Rrs)**:  
  - Rrs values equal to zero were excluded  
  - A derived feature, the Maximum Band Ratio (MBR), was computed as:  
    - Numerator: max(Rrs411, Rrs443, Rrs489, Rrs510)  
    - Denominator: Rrs555 + Rrs670  
  - MBR was log-transformed to create the primary input variable

- **Grouping and flags**:  
  - A categorical variable (`MBR_flag`) was used to indicate the dominant Rrs band used in the MBR numerator  
  - This grouping informed partial pooling structures in hierarchical models

### 2.2 Bayesian Modeling Framework

All models were implemented in PyMC v5 and estimated using Hamiltonian Monte Carlo with the No-U-Turn Sampler (NUTS). I modeled log-transformed chlorophyll-a as a function of log-transformed MBR, using a truncated normal likelihood to ensure predictions remained within physically plausible bounds.

Key components of the modeling framework include:

- **Likelihood**:  
  - Truncated normal with support between –3.0 and 3.2 (log₁₀ Chl)

- **Noise structure** (`σ`):  
  - Varied across models, with increasing structural complexity  
  - Some models use constant σ; others include heteroscedastic or hierarchical formulations

- **Priors**:  
  - Weakly informative priors were used throughout  
  - Regression coefficients received Normal priors  
  - All σ terms were modeled with Gamma distributions, allowing flexible but strictly positive support

- **Software**:  
  - PyMC for model specification and inference  
  - ArviZ for posterior diagnostics, posterior predictive checks, and model comparison

### 2.3 Overview of Candidate Models

I developed seven Bayesian models representing a progression of modeling complexity and assumptions:

- **Model 1**: Polynomial regression (Bayesian OC6-style baseline)  
- **Model 2**: Hierarchical linear regression, grouped by dominant MBR numerator band  
- **Model 3**: Hierarchical linear regression extended to include all Rrs spectral dominance groups  
- **Model 4**: Model 2 with group-specific constant σ  
- **Model 5**: Global linear heteroscedastic model  
- **Model 6**: Model 2 with group-wise linear heteroscedasticity  
- **Model 7**: Model 6 plus an added dispersion term for fluorescence-based measurements

The main text focuses on Models 1, 2, 6, and 7, which span the range of key structural variations. Full equations, priors, and diagnostic results for all models are provided in the Supplement.

### 2.4 Representative Model Structures

#### 2.4.1 Model 1: Bayesian Polynomial OC6 Regression

This model follows the OC6 formulation proposed by O’Reilly et al. (2019), adapted to a Bayesian framework. It uses a fourth-order polynomial in log(MBR) to estimate log(Chl).

- No grouping or partial pooling  
- Constant noise term σ shared across all observations  
- Gamma prior placed on σ

#### 2.4.2 Model 2: Hierarchical Linear Regression (4-band Grouping)

This model introduces partial pooling across four groups defined by the dominant MBR numerator band. Each group receives its own slope and intercept, with shared hyperpriors.

- Grouping: Rrs411, Rrs443, Rrs489, Rrs510  
- Partial pooling via shared hyperpriors  
- Constant σ across all groups

#### 2.4.3 Model 6: Hierarchical Linear Regression with Group-wise Linear σ

To accommodate heteroscedasticity, this model allows σ to vary with log(MBR) within each group, modeled as a log-linear relationship. Parameters are estimated hierarchically across groups.

- log(σ) = α + β × log(MBR), with group-specific α and β  
- Transformed via exp(log(σ))  
- Partial pooling of σ slope/intercept terms

#### 2.4.4 Model 7: Hierarchical Linear σ + Measurement-Specific Dispersion

This model adds a measurement-method-specific dispersion term to the structure of Model 6, capturing the known increase in variability associated with fluorescence-based chlorophyll estimates.

- Additional noise component applied when chl_flag = fluorescence  
- All other structure inherited from Model 6

### 2.5 Prior Specification and Inference Settings

I used weakly informative priors across all models to balance flexibility and regularization:

- **Regression terms**: Normal(0, 1) unless otherwise stated  
- **Hyperpriors**: Normal and HalfNormal for hierarchical parameters  
- **Noise terms** (`σ`): Gamma(γ, φ), chosen to reflect plausible dispersion values

Inference was conducted using:

- 4 chains per model  
- 1000–1500 tuning steps  
- ≥2000 posterior samples per chain

I assessed convergence using R-hat (target < 1.01) and effective sample size (ESS > 1000). Full trace plots and diagnostics are included in the Supplement.

### 2.6 Model Comparison and Predictive Evaluation

I compared model performance using **Pareto Smoothed Importance Sampling Leave-One-Out Cross-Validation (PSIS-LOO-CV)**, implemented via the `arviz.loo()` function. This Bayesian metric estimates the expected log predictive density (ELPD), a principled measure of out-of-sample predictive accuracy.

- Models were ranked by ELPD  
- ΔELPD values were interpreted relative to their standard errors  
- Pareto-k diagnostics confirmed LOO reliability (full table in Supplement)

I evaluated predictive accuracy and calibration using:

- **Posterior Predictive Checks (PPCs)**:  
  - Simulated draws overlaid on observed distributions  
  - Empirical CDF comparisons and interval coverage

- **HDI coverage**:  
  - Fraction of observed values falling within 95% posterior predictive intervals

No frequentist evaluation criteria such as RMSE, MAE, or R² were used at any stage as these do not make sense in this context where for each observations, a probability distribution of the prediction is produced.

## 3. Results

This section presents convergence diagnostics, posterior inference, and model comparison using fully Bayesian metrics. No frequentist evaluation criteria were used. Instead, we rely on information-theoretic model comparison (LOO-CV), posterior predictive checks, and visualization of uncertainty via highest density intervals (HDIs).

### 3.1 MCMC Convergence

All four models converged successfully. The Gelman-Rubin statistic (R-hat ≈ 1.0) and large effective sample sizes (ESS > 1000) across all parameters indicate robust posterior estimation. Diagnostic plots and sampling traces are provided in the Supplementary Material.

- All R-hat < 1.01
- No divergences
- Sufficient tail and bulk ESS
- Full diagnostics in Supplement

### 3.2 Predictive Performance via Leave-One-Out Cross Validation

We compared models using Pareto Smoothed Importance Sampling Leave-One-Out Cross-Validation (PSIS-LOO-CV), a fully Bayesian approach to evaluating out-of-sample predictive accuracy.

**Table 1.** Leave-One-Out Comparison (ELPD ± SE):

<!-- Insert Quarto table of ELPD, ΔELPD, SE, p_loo -->

- Model 7 ranked highest in expected log predictive density (ELPD)
- Model 6 performed comparably, with ΔELPD within 1 SE
- Model 1 (OC6 polynomial) had the weakest predictive performance
- All models had acceptable Pareto-k diagnostics (see Supplement)

### 3.3 Posterior Predictive Checks (PPC)

We used posterior predictive checks (PPCs) to evaluate the ability of each model to reproduce key features of the observed log-chlorophyll distribution. These include:

- Empirical CDF comparisons
- HDI envelopes overlaid on the data
- Predictive median vs. observed values

**Figure 1.** Posterior predictive envelopes vs. observed log(Chl)  
**Figure 2.** ECDF overlays with posterior predictive means and HDIs  
**Figure 3.** Rug plot of observations with overlaid posterior predictive draws

Findings:
- Model 7 captures full data distribution, especially for fluorescence-labeled values
- Model 1 underestimates tails (under-dispersed)
- Model 6 accurately reflects group-level variance

### 3.4 Posterior Distributions and Interpretability

Posterior distributions for key parameters illustrate how model structure affects inference. We show both:
- Group-level slopes and intercepts (hierarchical structure)
- Group-wise σ (heteroscedasticity)
- Fluorescence-specific variance term (Model 7)

**Figure 4.** Forest plot of slope and intercept posteriors  
**Figure 5.** Posterior of σ across MBR groups  
**Figure 6.** Posterior of added noise term for fluorescence

Key insights:
- Rrs510-dominant group had the steepest slope (Model 2 & 6)
- Fluorescence noise term is credibly nonzero (Model 7)
- Posterior variance structure supports heteroscedasticity as a key feature

### 3.5 Summary of Bayesian Comparison

- **Model 7**: Best predictive performance (highest ELPD), best uncertainty calibration
- **Model 6**: Nearly equivalent performance; simpler if fluorescence uncertainty not needed
- **Model 2**: Gains from partial pooling; underperforms without variance modeling
- **Model 1**: Legacy structure underfits tails, overconfident predictions

All results support the Bayesian hierarchical modeling framework with heteroscedasticity and measurement-aware noise as optimal for chlorophyll-a estimation.

## 4. Discussion

This section interprets the results in the context of previous chlorophyll-a modeling work, with particular emphasis on the added value of Bayesian modeling. We highlight the importance of accounting for heteroscedasticity and measurement-type uncertainty, and discuss how this contributes to improved predictive reliability and scientific interpretability.

### 4.1 Value of Bayesian Framework over Classical Approaches

Our results demonstrate the clear advantages of fully Bayesian modeling for chlorophyll-a prediction from satellite Rrs. Unlike deterministic or frequentist models:

- Bayesian methods explicitly model parameter uncertainty and provide HDIs instead of misleading confidence intervals
- Posterior predictive distributions enable direct evaluation of model calibration
- Bayesian model comparison via LOO-CV avoids overfitting and guards against inappropriate complexity

<!-- Expand: reinforce critique of RMSE-based selection; cite key Bayesian modeling references -->

### 4.2 Importance of Hierarchical Structuring

Models that leveraged group-wise structure (based on dominant Rrs bands) consistently outperformed simpler counterparts.

- Hierarchical partial pooling improved both prediction and parameter identifiability
- Group-level trends reflected known biogeophysical distinctions (e.g., clear vs. turbid water types)
- Intercepts and slopes varied systematically across spectral regimes

<!-- Expand: relate group-level results to known Case 1 vs. Case 2 water differences -->

### 4.3 Role of Heteroscedasticity in Chlorophyll Retrieval

Accounting for non-constant noise was essential for accurate uncertainty quantification.

- Models with linear or hierarchical σ terms captured increased variance at higher MBR or Chl
- Constant-σ models (e.g., OC6) underrepresented predictive uncertainty, especially in extremes
- Explicit modeling of σ structure produced calibrated HDIs and superior posterior predictive checks

<!-- Expand: link back to observed data behavior and prior studies showing variance scaling with concentration -->

### 4.4 Measurement-Specific Error Modeling

Model 7, which introduced a separate noise term for fluorescence-derived Chl, outperformed all others.

- Fluorescence-based measurements have higher and more variable error
- This heterogeneity cannot be ignored in merged datasets
- Bayesian modeling allows this variance to be encoded explicitly without discarding data

<!-- Expand: cite studies on fluorescence bias, show that Bayesian methods allow uncertainty to “absorb” method bias -->

### 4.5 Implications for Satellite Algorithm Development

The findings support incorporating Bayesian elements into future operational chlorophyll algorithms.

- Hierarchical modeling offers a path to generalize across bioregions and sensors
- Posterior uncertainty estimates can support downstream applications (e.g., forecasting, ecological thresholds)
- Bayesian frameworks are sensor-agnostic: Rrs inputs can vary as long as priors are adapted

<!-- Optional: comment on feasibility of deploying these methods, or emulating them with approximate Bayesian techniques -->

### 4.6 Limitations and Future Work

As with any modeling effort, our study has limitations:

- All results are conditional on the data from NOMAD; further validation on independent datasets is needed
- More sophisticated noise structures (e.g., nonparametric σ) could be explored
- Spatial or spatio-temporal extensions were not considered, but would be feasible in PyMC

<!-- Expand: suggest extensions (Gaussian Processes, spatial priors, stacking), and invite reproducibility -->

## Acknowledgments

## Open research

## References {.unnumbered}

:::{#refs}

:::