### 2025-05-21 manuscript/supplement.qmd
added supplementary materials and moved eqns there

---
title: "Supplementary Material"
subtitle: "Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation"
format:
  pdf:
    toc: true
    number-sections: true
    keep-tex: true
bibliography: references.bib
keep-tex: true
---

This document provides supplementary material supporting the manuscript titled _"Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation"_, submitted to *Remote Sensing of Environment*.

# Supplementary Methods

- Full Bayesian model specifications
- Prior predictive checks
- Diagnostic metrics for convergence (e.g., $\hat{R}$, ESS)
- Model comparison procedures (WAIC, LOO)

## Model 1 - Fourth Order Polynomial Regression
- model introduction (relation to Oreilly and Werdell 2019)

### Model Description and Priors
- brief word description

### Mathematical Description
\begin{align*}
    \alpha &\sim \text{Normal}(0, 1) \\
    \beta_1 &\sim \text{Normal}(0, 1) \\
    \beta_2 &\sim \text{Normal}(0, 1) \\
    \beta_3 &\sim \text{Normal}(0, 1) \\
    \beta_4 &\sim \text{Normal}(0, 1) \\
    \mu_i &= \alpha + \beta_1 X + \beta_2 X^2 + \beta_3 X^3 + \beta_4 X^4 \\
    \sigma &\sim \text{Gamma}(2, 2) \\
    y_i &\sim \text{TruncatedNormal}(\mu, \sigma, -3, 3.2)
\end{align*}

### Model Limitations
- Discussion of limitations (e.g., fixed structure, sensitivity to polynomial degree, difficulty of interpretation)

## Model 2: Hierarchical linear regression, grouped by dominant MBR numerator band

- model introduction (relation to Oreilly and Werdell 2019)

### Model Description and Priors
- brief word description

### Mathematical Description
- latex equations
### Model Limitations
- Discussion of limitations

## Model 3: Similar to Model 2, with group-specific constant likelihood variance  
- model introduction (relation to Oreilly and Werdell 2019)

### Model Description and Priors
- brief word description

### Mathematical Description
- latex equations
### Model Limitations
- Discussion of limitations

## Model 4: Global linear heteroscedastic model 
- model introduction (relation to Oreilly and Werdell 2019)

### Model Description and Priors
- brief word description

### Mathematical Description
- latex equations
### Model Limitations
- Discussion of limitations

## Model 5: Hierarchical Linear Regression with Linear Group-Wise Heteroscedasticity 
- model introduction

### Model Description and Priors
- brief word description of the idea behind the model and the priors used.

\begin{align*}
    \alpha_\mu &\sim \text{Normal}(0, 1) \\
    \alpha_\sigma &\sim \text{Exponential}(1) \\
    \beta_\mu &\sim \text{Normal}(0, 1) \\
    \beta_\sigma &\sim \text{Exponential}(1) \\
    \sigma_{\alpha_\mu} &\sim \text{Normal}(0, 1) \\
    \sigma_{\alpha_\sigma} &\sim \text{Exponential}(1) \\
    \sigma_{\beta_\mu} &\sim \text{Normal}(0, 1) \\
    \sigma_{\beta_\sigma} &\sim \text{Exponential}(1) \\
    \alpha_j &\sim \text{Normal}(\alpha_\mu, \alpha_\sigma) \quad \text{for } j = 1, \dots, J \\
    \beta_j &\sim \text{Normal}(\beta_\mu, \beta_\sigma) \quad \text{for } j = 1, \dots, J \\
    \sigma_{\alpha_j} &\sim \text{Normal}(\sigma_{\alpha_\mu}, \sigma_{\alpha_\sigma}) \quad \text{for } j = 1, \dots, J \\
    \sigma_{\beta_j} &\sim \text{Normal}(\sigma_{\beta_\mu}, \sigma_{\beta_\sigma}) \quad \text{for } j = 1, \dots, J \\
    \mu_i &= \alpha_{g_i} + \beta_{g_i} \cdot \log\text{MBR}_i \quad \text{for } i = 1, \dots, N \\
    \log \sigma_i &= \sigma_{\alpha_{g_i}} + \sigma_{\beta_{g_i}} \cdot \log\text{MBR}_i \\
    \sigma_i &= \exp(\log \sigma_i) \\
    \log\text{Chl}_i &\sim \text{TruncatedNormal}(\mu_i, \sigma_i, -3, 3.2)
\end{align*}

# Supplementary Figures

```{figure} figs/ppc_example.png
---
width: 85%
fig-cap: "Figure S1: Posterior predictive check comparing observed and simulated chlorophyll-a distributions across groups."
---