### 2025-05-21 manuscript/manuscript.qmd
moved eqns to Suppl. Updated model 5 in methods

---
title: "Shifting paradigms in Ocean Color: Bayesian Inference for Uncertainty-Aware Chlorophyll Estimation "
format:
    agu-pdf:
        keep-tex: true
    agu-html: default
author:  
  - name: Erdem M. Karaköylü
    orcid: 0000-0002-6156-1720
    corresponding: true
    email: erdemk@protonmail.com
    roles:
      - Investigator
      - Software
      - Visualization
      - Project administration
    affiliations:
      - Independant Researcher
  - name: Susanne E. Craig
    orcid: 0000-0002-8963-0951
    corresponding: false
    email: susanne.e.craig@nasa.gov
    roles:
      - Co-Investigator
    affiliations:
      - NASA/UMBC
abstract: Placeholder
plain-language-summary: Placeholder
keywords: ['Bayesian modeling', 'Ocean Color', ]
key-points:
  - Bayesian models have not been fully adopted in Ocean Color remote sensing yet.
  - These models are a principled way to take uncertainty into account
  - 
bibliography: references.bib
citation:
  container-title: Geophysical Research Letters
keep-tex: true
number-sections: true
date: last-modified
---

## Introduction

Satellite ocean color remote sensing has long served as a cornerstone of marine ecosystem monitoring, offering global and synoptic coverage of surface ocean properties. Among these, chlorophyll-a ($Chl_a$) concentration remains a central metric, widely used as a proxy for phytoplankton biomass, primary production, and water quality. The retrieval of $Chl_a$ from ocean color data has evolved over decades, resulting in a diverse lineage of empirical and semi-empirical algorithms. The following section summarizes this historical development, which sets the stage for a critical examination of the statistical foundations underlying current approaches.

### Historical Context of Chlorophyll Algorithms

Satellite ocean color observations have long been fundamental for monitoring marine ecosystems, as they enable global estimation of chlorophyll‑a ($Chl_a$) — a key indicator of phytoplankton biomass and ocean productivity. Early empirical algorithms, notably developed by O’Reilly et al. [@oreilly1998; @oreilly2000], established the $OCx$ family (where $x$ denotes the number of bands used) of polynomial regression models. These models relate blue-to-green reflectance ratios (after log‑transformation) to in situ $Chl_a$, employing either straight band ratios (BR) or maximum band ratios (MBR)—the latter selecting the highest available blue-to-green ratio for any given observation as input to a high‑order polynomial. These formulations have served as the operational foundation for chlorophyll‑a products across a broad range of satellite ocean color sensors—from the pioneering Coastal Zone Color Scanner (CZCS) through SeaWiFS, MODIS, and MERIS to more recent missions—offering a straightforward and robust approach for Case‑1 waters. However, their performance is more limited in optically complex Case‑2 waters and remains sensitive to atmospheric correction errors.

Subsequent refinements were introduced to address these deficiencies. For example, Hu et al. [@hu2012novel] proposed a Color Index (CI) formulation that employs a band‑difference approach to reduce sensitivity to residual atmospheric errors and instrument noise, with further improvements enhancing inter‑sensor agreement [@hu2019improving]. The increasing availability of calibration data (e.g., [@Valente2015]) and ongoing algorithmic improvements have led to the development of additional variants of the $OCx$ algorithms—specifically, the OC5 and OC6 formulations. O’Reilly and Werdell [@oreilly2019] maintain that OC5 extends the spectral basis by incorporating the 412 nm band, thereby exploiting its strong signal in clear, oligotrophic waters, while OC6 replaces the traditional denominator with the mean of the 555 and 670 nm reflectances, with the aim of improving the dynamic range at low chlorophyll concentrations. In total, [@oreilly2019] propose 65 versions of BR/MBR $OCx$ type algorithms for 25 sensors—on average, two or more variants per sensor. With this arsenal, it is hoped, researchers are better equipped to address the wide array of bio‑optical environments encountered in global ocean color applications.

### Limitations of Existing Approaches

Regrettably, the development of traditional ocean color algorithms is grounded in a fundamental statistical error—one that pervades much of observational science: the conflation of sampling probability with inferential probability [@jaynes2003probability; @DeScheemaekere2011].

Consider a dataset $D$ composed of input–output pairs—e.g., remote sensing reflectance (Rrs) and chlorophyll-a concentration ($Chl_a$)—and a model $M$, such as OCx, posited to represent the relationship between them. The sampling probability $p(D \mid M)$ denotes the probability of observing data $D$ under the assumption that model $M$ is true. In standard model fitting, this likelihood is maximized by adjusting the parameters of $M$ to best explain the observed data.

This approach tacitly assumes that the model which best fits the data also most accurately represents the underlying generative process. This constitutes an epistemic fallacy—treating $p(D \mid M)$ as if it were $p(M \mid D)$—a direct violation of Bayes’ theorem and the rules governing conditional probability.

Although in well-behaved, data-rich cases—where the likelihood is regular, the signal strong, and the model adequately constrained—the maxima of $p(D \mid M)$ and $p(M \mid D)$ may coincide, this remains the exception—not the rule.

This mistake lies at the heart of what Clayton [@clayton2022bernoulli] terms the Bernoulli Fallacy: the widespread misinterpretation of likelihood as inference, or of data-fit as belief. As Clayton argues, this logical misstep has far-reaching consequences, with implications that extend beyond science to domains such as medicine, law, and public policy.

In scientific modeling, this fallacy contributes to poor generalization, drives the use of ad hoc or retrospective uncertainty quantification, and underlies many published results that later prove difficult to replicate [@baker2016; @cobey2024biomedical]. These limitations are not restricted to classical hypothesis testing; they persist in the training and deployment of modern machine learning models as well.

In regression and classification, maximizing likelihood is often treated as sufficient for inference—despite yielding only a single point estimate and ignoring both parameter uncertainty and the plausibility of alternative models.

This epistemic shortcut has been directly critiqued in the machine learning literature. Gal [@gal2016uncertainty] and Ghahramani [@gahramani2015probabilistic] point out that most ML models discard uncertainty altogether, treating the outcome of an optimization as if it were an inference. The result is overconfident predictions and brittle generalization—concerns that echo Clayton’s critique.

Bishop [@bishop2006pattern] similarly distinguishes between the utility of predictive models and the inferential scaffolding required to quantify uncertainty, reinforcing the notion that likelihood alone is insufficient—and that the Bernoulli Fallacy permeates much of applied machine learning.

### Overcoming limitations

Several recent efforts have attempted to address the limitations of classical retrieval models. For instance, [@seegers2018] proposed alternative evaluation metrics to move beyond restrictive frequentist assumptions. Others have incorporated Bayesian elements into the modeling pipeline: [@frouin2013] applied Bayesian inversion for atmospheric correction, [@shi2015] used probabilistic fusion for multi-sensor data, and [@craig2019] employed Hamiltonian Monte Carlo to train Bayesian neural networks (BNNs) for retrieving inherent optical properties (IOPs) from top-of-atmosphere radiance. Similarly, [@werther2022] introduced Monte Carlo dropout as an approximate BNN strategy, while [@erickson2023] recast the Generalized Inherent Optical Property (GIOP) framework using conjugate Bayesian linear models. Most recently, [@hammout2024] developed a BNN surrogate using stochastic variational inference for chlorophyll-a prediction from Rrs.

While these studies mark important progress, they often adopt only isolated components of what has come to be known as the Bayesian workflow [@bgelman2019]. Key aspects—such as principled prior specification, posterior predictive checking, and formal model comparison—remain largely absent. As a result, uncertainty is frequently approximated rather than inferred, and model structure is rarely treated as a variable to be interrogated.

This paper builds on prior work and extends it by applying a fully Bayesian modeling framework for chlorophyll-a prediction from sea surface reflectance (Rrs), with explicit treatment of uncertainty, model complexity, and measurement error.

## 2. Materials and Methods

This section outlines the dataset, model structures, inference methods, and model evaluation criteria used in this study. I focus on a fully Bayesian approach to modeling chlorophyll-a from satellite remote sensing reflectance (Rrs), emphasizing transparency, uncertainty quantification, and principled model comparison. Four  models are presented in the main text, while all six model formulations and results are included in the Supplement.

### 2.1 Data Description and Preparation Overview

For pedagogical reasons, I used the well known NOMAD [@werdell2005improved] dataset, which includes quality-controlled in situ measurements of chlorophyll-a matched with coincident satellite Rrs data. These measurements span a range of oceanographic conditions, enabling the development of models that are expected to generalize across water types. The data were loaded using the Python Pandas library [@reback2020pandas], which was used for all subsequent cleaning, preprocessing, and transformation operations.

*Raw Data* - Sea-surface reflectances (Rrs) in the 6 visible SeaWiFS bands (411, 443, 489, 510, 550, 670nm) constitue the bulk of the data. Chlorophyll concentrations, measured via fluorescence or HPLC or both, consituted the target variables. Any row containing an invalid data point (null, zero, or negative) was removed, as proper missing/invalid data imputation is a subject onto itself and beyond the scope of this paper.

*Data Preprocessing and Transformation* - Chlorophyll measurements from the two different methods were combined into a single data. An auxiliary column was used to flag the measurement method as 'fluo' of 'hplc'. The chlorophyll was then log-transformed to stabilize its variance. The max band ratio (MBR) was computed as $\frac {\max(Rrs411, Rrs443, Rrs489, Rrs)}{Rrs555 + Rrs670}$, similar to the OC6 model proposed by [@oreilly2019]. Another ancillary column was created to captured the actual Rrs band used for each set observations. This numerator type as well as the chlorophyll measurement type are useful to build more hierarchical partial pooling models that can take advantage of such data groups. A sample from the fully processed data used in the subsequent models can be seen in [@data-table].

:::{#data-table}

|   | log_MBR              | MBR_flag   | log_chl              | hplc_flag   |
|----:|:-----------------------|:-----------|:-----------------------|:------------|
| 905 | -0.2545615131431814  | Rrs510     |  0.5837653682849998  | hplc        |
|  91 | -0.3641327253450254  | Rrs510     |  1.0511525224473812  | fluo        |
| 528 | -0.0737170739918758  | Rrs510     |  0.14363923527454328 | fluo        |
| 969 |  0.5813600497877703  | Rrs411     | -0.8272803386505035  | fluo        |
| 552 |  0.7134341855521412  | Rrs443     | -0.9318141382538383  | hplc        |

A random 5-row sample from the Pandas dataframe holding the data used in subsequent models. *log_MBR* is the log-transformed Max. Band Ratio (MBR) of sea surface reflectances, where the numerator is Max(Rrs411, Rrs443, Rrs489, Rrs510) and the denominator is computed from the sum of Rrs555 and Rrs670. *MBR_flag* indicates which of the 4 possible Rrs bands is used in the numerator for that particular row. *log_chl* is log transformed in-situ measured chlorophyll *a* concentration, originally in $mg\ m^{-3}$. *hplc_flag* is a binary variable indicating whether the measurement was obtained via HPLC or fluorometry. "-flag" are used during modeling for grouping data for hierarchical partial pooling models, explained further below. These groupings can be seen in [@fig-eda]

:::

::: {#fig-eda layout-ncol=2}

![](images/eda_numerator.png){#fig-eda-mbr}

![](images/eda_hplc.png){#fig-eda-hplc}

Two ways of grouping the data. [@fig-eda-mbr], max band ratio numerator; [@fig-eda-hplc], chlorophyll measurement method.
:::

### 2.2 Bayesian Modeling Framework

Several models of increasing levels were built.  All models use $log(MBR)$ as primary input and $log(chl)$ as target, some use $MBR\_flag$ as additional input, the last model also uses $hlpc_flag$. The following are key components shared across models.

*Likelihood* - A truncated gaussian distribution with support constrained to -3.0 to 3.2 on the $log_{10}$ scale,corresponding to $0.001 - 1600 mg\ m^{-3}$ to reflect both detectable realistic levels of marine surface chlorophyll.

*Noise Structure* - Noise was incorporated as likelihood dispersion parameter $σ$; some of the models use a constant $σ$, others take heteroscedasticity into account.

*Priors* - To estimate model posterior distribution, and produce predictions with uncertainty following data fitting, all model parameters receive a prior distribution, making explicit model assumptions. Often there will be enough data to overrule the priors, which are here mostly a conduit to the Posterior. An important caveat is that care should be taken not to assign a 0-probability through a poorly specified prior where support might actually be needed. Less importantly, more efficient use of the search space can be done through priors that assign 0-probability where support is clearly unreasonable. Thus, all regression parameters receive Normal (Gaussian) 0-centered weakly informative priors. And all $σ$ terms received Gamma priors that provide flexible but strictly positive support.

*Software* -  All models were implemented with the Python probabilistic programming library PyMC V.5 [@abril2023pymc] and estimated using a variant of Hamiltonian Monte Carlo known as the No-U-Turn Sampler (NUTS). Model diagnostics inference evaluation and model comparisons were done through the Arviz package.

###  Overview of Candidate Models

I developed six models representing a progression of modeling complexity and assumptions:

- **Model 1**: Polynomial regression (Bayesian OC6-style baseline)  
- **Model 2**: Hierarchical linear regression, grouped by dominant MBR numerator band  
- **Model 3**: Similart to Model 2, with group-specific constant likelihood variance  
- **Model 4**: Global linear heteroscedastic model  
- **Model 5**: Similar to Model 2, with group-wise linear heteroscedasticity  
- **Model 6**: Model 6 plus an added dispersion term for fluorescence-based measurements

The main text focuses on Models 1, 2, 5, and 6, which span the range of key structural variations. Full equations, priors, and diagnostic results for all models are provided in the Supplement.

### Representative Model Structures
Models are described here as sets of equations and Directed Acyclic Graphs (DAG), with the former providing numerical details while the latters helps visualize model structure and flow.

#### Model 1: Bayesian 4th Order Polynomia Regression

This model follows the OC6 formulation proposed by O’Reilly et al. (2019), with the difference that the denominator is the sum of $Rrs555$ and $Rrs670$ rather than the mean. The model mean is a fourth-order polynomial of $log(MBR)$ to estimate $log(Chl)$. The set of proabilistic equations that define this model can be found in the Supplemental Material Section.
This model can be represented by the following set of probabilistic equations where $X$ represents $log(MBR)$, and $y_i$ correspondonds to the likelihood of the $i^{th}$ $log(Chl)$ observation

::: {#fig-oc6-struct}

![](images/oc6_structure.png)

Model 1 DAG showing model parameters and distributions used as priors. The intercept $α$ and the polynomial slopes, $β1$-$β4$ all have Gaussian priors; $σ$, the likelihood dispersion parameter, receives a Gamma prior; the likelihood itself is constrained by a Truncated Normal distribution (see text for more). The plate containing the likelihood is indexed ($obs\_idx$) indicating the likelihood will be computed for each of the 1082 observations contained in the dataframe and show the number of data points available for fitting.

:::

#### Model 2: Hierarchical Linear Regression (4-band Grouping)

*Model 2* introduces partial pooling across four groups defined by the dominant MBR numerator band; Rrs411, Rrs443, Rrs489, Rrs510. Each group receives its own slope and intercept with shared hyperpriors. On the one hand each group has its own submodel that can capture details specific to it, while the shared hyperpriors allow for the sharing of information between groups when fitting the model, thus maximizing the use of information contained in the available data. $X_i$ is the same as for Model 1, i.e. $log(MBR)$, assigned a given observation to group $i$ (@fig-eda-mbr). This model retains a constant dispersion parameter $σ$, similar to *Model 1*.
Below is the system of equations corresponding to this model, where $j$ is the group index and $i$ is the $i^{th}$ obasrvation. Mathematical description of the model can be found in the Supplementary Material. @fig-model2-struct illustrates the model structure.

::: {#fig-model2-struct}

![](images/model2_structure.png)

Model 2 - Multi-level hierarchical level with partial pooling across $MBR$ numerator groups (Rrs411, Rrs443, Rrs489, Rrs510). Model parameter prior mean and variance have Gaussian and Exponential hyperpriors. This structure allows the sharing of information across groups at model fitting time.
:::

#### Model 5: Hierarchical Linear Regression with Linear Heteroscedasticity

The data plotted in @fig-eda-mbr suggests heteroscedasticity - variance that is not constant. Morevoer variance seems to change with the data. Models 3 and 4 (see Supplemental Material section) investigated the role of groupwise constant heteroscedasticity and global linear heteroscedasticity, respectively. To investigate this further, Model 5 include groupwise linear heteroscedasticity. Thus,  for a given MBR group $j$ $σ_j$ is modeled as a log-linear relationship such that $$log(σ) = α_j + β_j × log(MBR)$ and then transformed via $exp(log(σ_j))$ before being passed to the likelihood. Similarly to the model mean $μ_j$, $σ_j$'s slope and intercept terms are partially pooled and so have common hyperpriors. Mathematical descriiption of the model can be found in the Supplementary Material section; @fig-model5-struct.  illustrates the model structure.

::: {#fig-model5-struct}

![](images/model5_structure.png)

Model 5 - Multi-level hierarchical linear model with partial pooling across $MBR$ numerator groups (Rrs411, Rrs443, Rrs489, Rrs510) for both model mean and model variance.  Model parameter prior mean and variance have Gaussian and Exponential hyperpriors similar to Model 2s.
:::

####  Model 6:  Measurement-Specific Dispersion

This model adds a measurement-method-specific dispersion term to the structure of Model 6, capturing the known increase in variability associated with fluorescence-based chlorophyll estimates.

- Additional noise component applied when chl_flag = fluorescence  
- All other structure inherited from Model 6

### Inference Settings

I used weakly informative priors across all models to balance flexibility and regularization:

- **Regression terms**: Normal(0, 1) unless otherwise stated  
- **Hyperpriors**: Normal and HalfNormal for hierarchical parameters  
- **Noise terms** (`σ`): Gamma(γ, φ), chosen to reflect plausible dispersion values

Inference was conducted using:

- 4 chains per model  
- 1000–1500 tuning steps  
- ≥2000 posterior samples per chain

I assessed convergence using R-hat (target < 1.01) and effective sample size (ESS > 1000). Full trace plots and diagnostics are included in the Supplement.

### Model Comparison and Predictive Evaluation

I compared model performance using **Pareto Smoothed Importance Sampling Leave-One-Out Cross-Validation (PSIS-LOO-CV)**, implemented via the `arviz.loo()` function. This Bayesian metric estimates the expected log predictive density (ELPD), a principled measure of out-of-sample predictive accuracy.

- Models were ranked by ELPD  
- ΔELPD values were interpreted relative to their standard errors  
- Pareto-k diagnostics confirmed LOO reliability (full table in Supplement)

I evaluated predictive accuracy and calibration using:

- **Posterior Predictive Checks (PPCs)**:  
  - Simulated draws overlaid on observed distributions  
  - Empirical CDF comparisons and interval coverage

- **HDI coverage**:  
  - Fraction of observed values falling within 95% posterior predictive intervals

No frequentist evaluation criteria such as RMSE, MAE, or R² were used at any stage as these do not make sense in this context where for each observations, a probability distribution of the prediction is produced.

## Results

This section presents the outputs of posterior inference, model comparison using Bayesian criteria, and posterior predictive diagnostics for the four primary models. Results from the three supplementary models are provided separately in the Supplement.

### MCMC Convergence

All four models converged successfully. The Gelman-Rubin statistic (R-hat ≈ 1.0) and large effective sample sizes (ESS > 1000) across all parameters indicate robust posterior estimation. Diagnostic plots and sampling traces are provided in the Supplementary Material.

- All R-hat < 1.01
- No divergences
- Sufficient tail and bulk ESS
- Full diagnostics in Supplement

### Predictive Performance via Leave-One-Out Cross Validation

We compared models using Pareto Smoothed Importance Sampling Leave-One-Out Cross-Validation (PSIS-LOO-CV), a fully Bayesian approach to evaluating out-of-sample predictive accuracy.

**Table 1.** Leave-One-Out Comparison (ELPD ± SE):

<!-- Insert Quarto table of ELPD, ΔELPD, SE, p_loo -->

- Model 7 ranked highest in expected log predictive density (ELPD)
- Model 6 performed comparably, with ΔELPD within 1 SE
- Model 1 (OC6 polynomial) had the weakest predictive performance
- All models had acceptable Pareto-k diagnostics (see Supplement)

### 3.3 Posterior Predictive Checks (PPC)

We used posterior predictive checks (PPCs) to evaluate the ability of each model to reproduce key features of the observed log-chlorophyll distribution. These include:

- Empirical CDF comparisons
- HDI envelopes overlaid on the data
- Predictive median vs. observed values

**Figure 1.** Posterior predictive envelopes vs. observed log(Chl)  
**Figure 2.** ECDF overlays with posterior predictive means and HDIs  
**Figure 3.** Rug plot of observations with overlaid posterior predictive draws

Findings:
- Model 7 captures full data distribution, especially for fluorescence-labeled values
- Model 1 underestimates tails (under-dispersed)
- Model 6 accurately reflects group-level variance

### 3.4 Posterior Distributions and Interpretability

Posterior distributions for key parameters illustrate how model structure affects inference. We show both:
- Group-level slopes and intercepts (hierarchical structure)
- Group-wise σ (heteroscedasticity)
- Fluorescence-specific variance term (Model 7)

**Figure 4.** Forest plot of slope and intercept posteriors  
**Figure 5.** Posterior of σ across MBR groups  
**Figure 6.** Posterior of added noise term for fluorescence

Key insights:
- Rrs510-dominant group had the steepest slope (Model 2 & 6)
- Fluorescence noise term is credibly nonzero (Model 7)
- Posterior variance structure supports heteroscedasticity as a key feature

### 3.5 Summary of Bayesian Comparison

- **Model 7**: Best predictive performance (highest ELPD), best uncertainty calibration
- **Model 6**: Nearly equivalent performance; simpler if fluorescence uncertainty not needed
- **Model 2**: Gains from partial pooling; underperforms without variance modeling
- **Model 1**: Legacy structure underfits tails, overconfident predictions

All results support the Bayesian hierarchical modeling framework with heteroscedasticity and measurement-aware noise as optimal for chlorophyll-a estimation.

## 4. Discussion

This section interprets the results in the context of previous chlorophyll-a modeling work, with particular emphasis on the added value of Bayesian modeling. We highlight the importance of accounting for heteroscedasticity and measurement-type uncertainty, and discuss how this contributes to improved predictive reliability and scientific interpretability.

### 4.1 Value of Bayesian Framework over Classical Approaches

Our results demonstrate the clear advantages of fully Bayesian modeling for chlorophyll-a prediction from satellite Rrs. Unlike deterministic or frequentist models:

- Bayesian methods explicitly model parameter uncertainty and provide HDIs instead of misleading confidence intervals
- Posterior predictive distributions enable direct evaluation of model calibration
- Bayesian model comparison via LOO-CV avoids overfitting and guards against inappropriate complexity

<!-- Expand: reinforce critique of RMSE-based selection; cite key Bayesian modeling references -->

### 4.2 Importance of Hierarchical Structuring

Models that leveraged group-wise structure (based on dominant Rrs bands) consistently outperformed simpler counterparts.

- Hierarchical partial pooling improved both prediction and parameter identifiability
- Group-level trends reflected known biogeophysical distinctions (e.g., clear vs. turbid water types)
- Intercepts and slopes varied systematically across spectral regimes

<!-- Expand: relate group-level results to known Case 1 vs. Case 2 water differences -->

### 4.3 Role of Heteroscedasticity in Chlorophyll Retrieval

Accounting for non-constant noise was essential for accurate uncertainty quantification.

- Models with linear or hierarchical σ terms captured increased variance at higher MBR or Chl
- Constant-σ models (e.g., OC6) underrepresented predictive uncertainty, especially in extremes
- Explicit modeling of σ structure produced calibrated HDIs and superior posterior predictive checks

<!-- Expand: link back to observed data behavior and prior studies showing variance scaling with concentration -->

### 4.4 Measurement-Specific Error Modeling

Model 7, which introduced a separate noise term for fluorescence-derived Chl, outperformed all others.

- Fluorescence-based measurements have higher and more variable error
- This heterogeneity cannot be ignored in merged datasets
- Bayesian modeling allows this variance to be encoded explicitly without discarding data

<!-- Expand: cite studies on fluorescence bias, show that Bayesian methods allow uncertainty to “absorb” method bias -->

### 4.5 Implications for Satellite Algorithm Development

The findings support incorporating Bayesian elements into future operational chlorophyll algorithms.

- Hierarchical modeling offers a path to generalize across bioregions and sensors
- Posterior uncertainty estimates can support downstream applications (e.g., forecasting, ecological thresholds)
- Bayesian frameworks are sensor-agnostic: Rrs inputs can vary as long as priors are adapted

<!-- Optional: comment on feasibility of deploying these methods, or emulating them with approximate Bayesian techniques -->

### 4.6 Limitations and Future Work

As with any modeling effort, our study has limitations:

- All results are conditional on the data from NOMAD; further validation on independent datasets is needed
- More sophisticated noise structures (e.g., nonparametric σ) could be explored
- Spatial or spatio-temporal extensions were not considered, but would be feasible in PyMC

<!-- Expand: suggest extensions (Gaussian Processes, spatial priors, stacking), and invite reproducibility -->

## Acknowledgments

## Open research

## References {.unnumbered}

:::{#refs}

:::